---
title: "Decorrelation-Based Feature Discovery: Glaucoma"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
  word_document: 
    reference_docx: WordStyle_FRESA.docx
    toc: yes
    fig_caption: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message = FALSE,comment = "#>")

```

# Effect of UPSTM-Based Decorrelation on Feature Discovery



### Loading the libraries

```{r}
library("FRESA.CAD")
library(readxl)
library(igraph)
library(umap)
library(tsne)
library(entropy)
library(TH.data)


op <- par(no.readonly = TRUE)
pander::panderOptions('digits', 3)
pander::panderOptions('table.split.table', 400)
pander::panderOptions('keep.trailing.zeros',TRUE)
source("C:/Users/jtame/Documents/GitHub/LatentBiomarkers/RMD/DecorrelationCV.R")

```



All variables are derived from a laser scanning image of the eye background taken by the Heidelberg Retina Tomograph. Most of the variables describe either the area or volume in certain parts of the papilla and are measured in four sectors (temporal, superior, nasal and inferior) as well as for the whole papilla (global). The global measurement is, roughly, the sum of the measurements taken in the four sector.

The observations in both groups are matched by age and sex to prevent any bias.

Source
Torsten Hothorn and Berthold Lausen (2003), Double-Bagging: Combining classifiers by bootstrap aggregation. Pattern Recognition, 36(6), 1303–1309.



## The Data


GlaucomaM {TH.data}

```{r}

data("GlaucomaM")

pander::pander(table(GlaucomaM$Class))

GlaucomaM$Class <- 1*(GlaucomaM$Class=="glaucoma")

```


#### Standarize the names for the reporting

```{r results = "asis"}
studyName <- "Glaucoma"
dataframe <- GlaucomaM
outcome <- "Class"



nLoops <- 100

needScaling <- TRUE
IDSample <- FALSE
thro <- 0.8
#cmethod="fast"

```

## CV Analysis and report

```{r parent, child = here::here("~/GitHub/LatentBiomarkers/RMD/Child_univariate/CV_analysis.Rmd")}
```


## Testing different parameters
```{r}

cvData_0.9 <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.9)
cvData_0.8 <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.8)
cvData_0.7 <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.7)
cvData_0.6 <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.6)
cvData_0.5 <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.5)
bpDe <- predictionStats_binary(cvData_0.9$DeMedTestRaw,"IDeA@0.9")
bpDe <- predictionStats_binary(cvData_0.8$DeMedTestRaw,"IDeA@0.8")
bpDe <- predictionStats_binary(cvData_0.7$DeMedTestRaw,"IDeA@0.7")
bpDe <- predictionStats_binary(cvData_0.6$DeMedTestRaw,"IDeA@0.6")
bpDe <- predictionStats_binary(cvData_0.5$DeMedTestRaw,"IDeA@0.5")

cvData_0.9_R <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.9,corRank=TRUE)
cvData_0.8_R <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.8,corRank=TRUE)
cvData_0.7_R <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.7,corRank=TRUE)
cvData_0.6_R <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.6,corRank=TRUE)
cvData_0.5_R <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.5,corRank=TRUE)
bpDe <- predictionStats_binary(cvData_0.9_R$DeMedTestRaw,"IDeA@0.9 CorRank")
bpDe <- predictionStats_binary(cvData_0.8_R$DeMedTestRaw,"IDeA@0.8 CorRank")
bpDe <- predictionStats_binary(cvData_0.7_R$DeMedTestRaw,"IDeA@0.7 CorRank")
bpDe <- predictionStats_binary(cvData_0.6_R$DeMedTestRaw,"IDeA@0.6 CorRank")
bpDe <- predictionStats_binary(cvData_0.5_R$DeMedTestRaw,"IDeA@0.5 CorRank")


cvData_0.9_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.9,Outcome=outcome)
cvData_0.8_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.8,Outcome=outcome)
cvData_0.7_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.7,Outcome=outcome)
cvData_0.6_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.6,Outcome=outcome)
cvData_0.5_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.5,Outcome=outcome)
bpDe <- predictionStats_binary(cvData_0.9_D$DeMedTestRaw,"IDeA@0.9 Driven")
bpDe <- predictionStats_binary(cvData_0.8_D$DeMedTestRaw,"IDeA@0.8 Driven")
bpDe <- predictionStats_binary(cvData_0.7_D$DeMedTestRaw,"IDeA@0.7 Driven")
bpDe <- predictionStats_binary(cvData_0.6_D$DeMedTestRaw,"IDeA@0.6 Driven")
bpDe <- predictionStats_binary(cvData_0.5_D$DeMedTestRaw,"IDeA@0.5 Driven")

cvData_0.9_S <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.9,method="spearman")
cvData_0.8_S <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.8,method="spearman")
cvData_0.7_S <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.7,method="spearman")
cvData_0.6_S <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.6,method="spearman")
cvData_0.5_S <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.5,method="spearman")
bpDe <- predictionStats_binary(cvData_0.9_S$DeMedTestRaw,"IDeA@0.9 Spearman")
bpDe <- predictionStats_binary(cvData_0.8_S$DeMedTestRaw,"IDeA@0.8 Spearman")
bpDe <- predictionStats_binary(cvData_0.7_S$DeMedTestRaw,"IDeA@0.7 Spearman")
bpDe <- predictionStats_binary(cvData_0.6_S$DeMedTestRaw,"IDeA@0.6 Spearman")
bpDe <- predictionStats_binary(cvData_0.5_S$DeMedTestRaw,"IDeA@0.5 Spearman")

cvData_0.9_S_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                         DecorPCAEFA=FALSE,thr=0.9,Outcome=outcome,method="spearman")
cvData_0.8_S_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                         DecorPCAEFA=FALSE,thr=0.8,Outcome=outcome,method="spearman")
cvData_0.7_S_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                         DecorPCAEFA=FALSE,thr=0.7,Outcome=outcome,method="spearman")
cvData_0.6_S_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                         DecorPCAEFA=FALSE,thr=0.6,Outcome=outcome,method="spearman")
cvData_0.5_S_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                         DecorPCAEFA=FALSE,thr=0.5,Outcome=outcome,method="spearman")
bpDe <- predictionStats_binary(cvData_0.9_S_D$DeMedTestRaw,"IDeA@0.9 Spearman Driven")
bpDe <- predictionStats_binary(cvData_0.8_S_D$DeMedTestRaw,"IDeA@0.8 Spearman Driven")
bpDe <- predictionStats_binary(cvData_0.7_S_D$DeMedTestRaw,"IDeA@0.7 Spearman Driven")
bpDe <- predictionStats_binary(cvData_0.6_S_D$DeMedTestRaw,"IDeA@0.6 Spearman Driven")
bpDe <- predictionStats_binary(cvData_0.5_S_D$DeMedTestRaw,"IDeA@0.5 Spearman Driven")


save.image(paste("ALLCV_analyiss",studyName,".RData",sep="_"))

```

