---
title: "PCA EFA: Parkinson's"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
  word_document: 
    reference_docx: WordStyle_FRESA.docx
    toc: yes
    fig_caption: yes
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message = FALSE,comment = "#>")
```

# Parkinson's PCA-EFA-GDSTM

### Loading the libraries

```{r}
library("FRESA.CAD")
library(psych)
library(whitening)

library(readxl)
op <- par(no.readonly = TRUE)
pander::panderOptions('digits', 3)
pander::panderOptions('table.split.table', 400)
pander::panderOptions('keep.trailing.zeros',TRUE)

```


## The Data

```{r}

pd_speech_features <- as.data.frame(read_excel("~/GitHub/FCA/Data/pd_speech_features.xlsx",sheet = "pd_speech_features", range = "A2:ACB758"))




```

### The Average of the Three Repetitions

Each subject had three repeated observations. Here I'll use the average of the three experiments per subject.

```{r results = "asis", warning = FALSE, dpi=600, fig.height= 6.0, fig.width= 8.0}
rep1Parkison <- subset(pd_speech_features,RID==1)
rownames(rep1Parkison) <- rep1Parkison$id
rep1Parkison$id <- NULL
rep1Parkison$RID <- NULL
rep1Parkison[,1:ncol(rep1Parkison)] <- sapply(rep1Parkison,as.numeric)

rep2Parkison <- subset(pd_speech_features,RID==2)
rownames(rep2Parkison) <- rep2Parkison$id
rep2Parkison$id <- NULL
rep2Parkison$RID <- NULL
rep2Parkison[,1:ncol(rep2Parkison)] <- sapply(rep2Parkison,as.numeric)

rep3Parkison <- subset(pd_speech_features,RID==3)
rownames(rep3Parkison) <- rep3Parkison$id
rep3Parkison$id <- NULL
rep3Parkison$RID <- NULL
rep3Parkison[,1:ncol(rep3Parkison)] <- sapply(rep3Parkison,as.numeric)

whof <- !(colnames(rep1Parkison) %in% c("gender","class"));
avgParkison <- rep1Parkison;
avgParkison[,whof] <- (rep1Parkison[,whof] + rep2Parkison[,whof] + rep3Parkison[,whof])/3
## I apply the log transform to the data
##avgParkison[,whof] <- signedlog(avgParkison[,whof])
#avgParkison[,whof] <- FRESAScale(avgParkison[,whof],method="OrderLogit")$scaledData
#pander::pander(table(avgParkison$class))

```

#### Standarize the names for the reporting

```{r results = "asis"}
dataframe <- avgParkison
outcome <- "class"

trainFraction <- 0.6
rhoThreshold <- 0.6
TopVariables <- 5

trainSample <- sample(nrow(dataframe),nrow(dataframe)*trainFraction)

trainDataFrame <- dataframe[trainSample,]
testDataFrame <- dataframe[-trainSample,]

```



```{r results = "asis"}
pander::pander(c(rows=nrow(dataframe),col=ncol(dataframe)-1))
pander::pander(table(dataframe[,outcome]))

varlist <- colnames(dataframe)
varlist <- varlist[varlist != outcome]
varlist <- as.data.frame(cbind(name=varlist,desc=varlist))

```


## Univariate

```{r results = "asis"}

univariate_columns <- c("caseMean","caseStd","controlMean","controlStd","controlKSP","ROCAUC","wilcox.Zvalue")
univar <- uniRankVar(varlist,
	           paste(outcome,"~1"),
	           outcome,
	           testDataFrame,
              rankingTest = "AUC")



```

## Decorrelation with GDSTM Blind

```{r results = "asis"}
DEdataframe <- GDSTMDecorrelation(trainDataFrame,thr=rhoThreshold,verbose = TRUE)
varlistDe <-  colnames(DEdataframe)[colnames(DEdataframe) != outcome];
varlistDe <- as.data.frame(cbind(name=varlistDe,desc=varlistDe))

pander::pander(head(getLatentCoefficients(DEdataframe)))
predTestDe <- predictDecorrelate(DEdataframe,testDataFrame)

univarDe <- uniRankVar(varlistDe,
              paste(outcome,"~1"),
	            outcome,
              predTestDe,
              rankingTest = "AUC")


```

## Decorrelation with GDSTM Driven

```{r results = "asis"}

DriDEdataframe <- GDSTMDecorrelation(trainDataFrame,Outcome=outcome,thr=rhoThreshold)
varlistDe <-  colnames(DriDEdataframe)[colnames(DriDEdataframe) != outcome];
varlistDe <- as.data.frame(cbind(name=varlistDe,desc=varlistDe))

pander::pander(head(getLatentCoefficients(DriDEdataframe)))
predTestDri <- predictDecorrelate(DriDEdataframe,testDataFrame)

univarDeDri <- uniRankVar(varlistDe,
              paste(outcome,"~1"),
	            outcome,
              predTestDri,
              rankingTest = "AUC")


```

## Decorrelation with GDSTM Driven and Spearman

```{r results = "asis"}

DriDEdataframeSpear <- GDSTMDecorrelation(trainDataFrame,Outcome=outcome,thr=rhoThreshold,method="spearman",verbose=TRUE)

pander::pander(head(getLatentCoefficients(DriDEdataframeSpear)))
predTestDriSpear <- predictDecorrelate(DriDEdataframeSpear,testDataFrame)

varlistDeSpear <-  colnames(DriDEdataframeSpear)[colnames(DriDEdataframeSpear) != outcome];
univarDeDriSpear <- uniRankVar(varlistDeSpear,
              paste(outcome,"~1"),
	            outcome,
              predTestDriSpear,
              rankingTest = "AUC")

```


## PCA Analysis

```{r results = "asis"}

### PCA 
noclassData <- trainDataFrame[,!(colnames(trainDataFrame) %in% c(outcome,"gender"))]
cmat <- cor(noclassData);
diag(cmat) <- 0;
maxcor <- apply(cmat,2,max);
topcor <- names(maxcor[maxcor>rhoThreshold]) ## Only correlaed features will be PCA
pander::pander(c(Ncor=length(topcor)))

notcorr <- colnames(trainDataFrame)[!(colnames(trainDataFrame) %in% topcor)]
noclassData <- noclassData[,topcor]
noclassTestData <- testDataFrame[,topcor]
pc <- principal(noclassData,2*TopVariables,rotate="varimax")   #principal components
pander::pander(head(pc$loadings))
PCA_Train <- as.data.frame(cbind(predict(pc,noclassData),trainDataFrame[,notcorr]))
PCA_Predicted <- as.data.frame(cbind(predict(pc,noclassTestData),testDataFrame[,notcorr]))

varlistPCA <-  colnames(PCA_Predicted)[colnames(PCA_Predicted) != "class"];
univarPCA <- uniRankVar(varlistPCA,
              paste(outcome,"~1"),
	            outcome,
              PCA_Predicted,
              rankingTest = "AUC")



```

## EFA
```{r results = "asis"}

uls <- fa(noclassData,2*TopVariables,rotate="varimax")  #unweighted least squares is minres 
pander::pander(head(uls$weights)) 
EFA_Train <- as.data.frame(cbind(predict(uls,noclassData),trainDataFrame[,notcorr]))
EFA_Predicted <- as.data.frame(cbind(predict(uls,noclassTestData),testDataFrame[,notcorr]))

varlistEFA <-  colnames(EFA_Predicted)[colnames(EFA_Predicted) != "class"];
univarEFA <- uniRankVar(varlistEFA,
              paste(outcome,"~1"),
	            outcome,
              EFA_Predicted,
              rankingTest = "AUC")


```


## White
```{r results = "asis"}
library(whitening)
WhiteMat = whiteningMatrix(cov(noclassData), method="PCA-cor")
WhiteMat[is.na(WhiteMat)] <- 0;

pander::pander(head(WhiteMat[,c(1:TopVariables)])) 
PCAWhite_Train <- as.data.frame(cbind(tcrossprod(as.matrix(noclassData), WhiteMat),trainDataFrame[,notcorr]))
PCAWhitePredicted <- as.data.frame(cbind(tcrossprod(as.matrix(noclassTestData), WhiteMat),testDataFrame[,notcorr]))

varlistWhite <-  colnames(PCAWhitePredicted)[colnames(PCAWhitePredicted) != "class"];
univarWhite <- uniRankVar(varlistWhite,
              paste(outcome,"~1"),
	            outcome,
              PCAWhitePredicted,
              rankingTest = "AUC")


```

## Correlation Matrices

### RAW

```{r results = "asis", warning = FALSE, dpi=600, fig.height= 5, fig.width= 6}
par(mfrow=c(1,1),cex=0.6,cex.main=0.8)
breaks <- c(0:5)/5.0;

cormat <- cor(testDataFrame,method="spearman")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;
pander::pander(max(abs(cormat)))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(Raw_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

gplots::heatmap.2(abs(cormat),
                  trace = "none",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  breaks = breaks,
                  main = "Raw Correlation",
                  cexRow = 0.25,
                  cexCol = 0.25,
                  srtCol=35,
                  srtRow=75,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature"
                  )

hist(cormat,freq=FALSE,
     density=NULL,
     xlim=c(-1,1),
     ylim=c(0,4.0),
     main="Raw Correlation",xlab="Spearman Correlation")


```

### GDSTM Blind

```{r results = "asis", warning = FALSE, dpi=600, fig.height= 5, fig.width= 6}
par(mfrow=c(1,1),cex=0.6,cex.main=0.8)

## Train Correlation

cormat <- cor(DEdataframe,method="pearson")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

#pander::pander(colnames(cormat)[apply(abs(cormat),2,max)>rhoThreshold])

pander::pander(c(Train=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(HMCA_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

## Test Correlation
cormat <- cor(predTestDe,method="spearman")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

pander::pander(c(Test=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(HMCA_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

gplots::heatmap.2(abs(cormat),
                  trace = "none",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  breaks = breaks,
                  main = "Correlation after HMCA",
                  cexRow = 0.25,
                  cexCol = 0.25,
                  srtCol=35,
                  srtRow=75,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")

hist(cormat,freq=FALSE,
     density=NULL,
     xlim=c(-1,1),
     ylim=c(0,4.0),
     main="Correlation after GDSTM",xlab="Spearman Correlation")


```

### GDSTM Driven

```{r results = "asis", warning = FALSE, dpi=600, fig.height= 5, fig.width= 6}
par(mfrow=c(1,1),cex=0.6,cex.main=0.8)

## Train Correlation

cormat <- cor(DriDEdataframe,method="pearson")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;
#pander::pander(colnames(cormat)[apply(abs(cormat),2,max)>rhoThreshold])

pander::pander(c(Train=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(HMCA_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

## Test Correlation
cormat <- cor(DriDEdataframe,method="spearman")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

pander::pander(c(Test=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(HMCA_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

gplots::heatmap.2(abs(cormat),
                  trace = "none",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  breaks = breaks,
                  main = "Correlation after Driven-HMCA",
                  cexRow = 0.25,
                  cexCol = 0.25,
                  srtCol=35,
                  srtRow=75,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")
hist(cormat,freq=FALSE,
     density=NULL,
     xlim=c(-1,1),
     ylim=c(0,4.0),
     main="Correlation after Driven-GDSTM",xlab="Spearman Correlation")


```
### GDSTM Spearman
```{r results = "asis", warning = FALSE, dpi=600, fig.height= 5, fig.width= 6}
par(mfrow=c(1,1),cex=0.6,cex.main=0.8)

## Train Correlation

cormat <- cor(DriDEdataframeSpear,method="pearson")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;
#pander::pander(colnames(cormat)[apply(abs(cormat),2,max)>rhoThreshold])

pander::pander(c(Train=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(HMCAS_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

## Test Correlation

cormat <- cor(predTestDriSpear,method="spearman")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

pander::pander(c(Test=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(HMCAS_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

gplots::heatmap.2(abs(cormat),
                  trace = "none",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  breaks = breaks,
                  main = "Correlation after HMCA/Spearman",
                  cexRow = 0.25,
                  cexCol = 0.25,
                  srtCol=35,
                  srtRow=75,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")
hist(cormat,freq=FALSE,
     density=NULL,
     xlim=c(-1,1),
     ylim=c(0,4.0),
     main="Correlation after GDSTM with Spearman",xlab="Spearman Correlation")

```

### PCA

```{r results = "asis", warning = FALSE, dpi=600, fig.height= 5, fig.width= 6}
par(mfrow=c(1,1),cex=0.6,cex.main=0.8)



## Train Correlation

cormat <- cor(PCA_Train,method="pearson")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

pander::pander(c(Train=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(PCA_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

## Test Correlation
cormat <- cor(PCA_Predicted,method="spearman")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

pander::pander(c(Test=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(PCA_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

gplots::heatmap.2(abs(cormat),
                  trace = "none",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  breaks = breaks,
                  main = "Correlation after PCA",
                  cexRow = 0.25,
                  cexCol = 0.25,
                  srtCol=35,
                  srtRow=75,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")


hist(cormat,freq=FALSE,
     density=NULL,
     xlim=c(-1,1),
     ylim=c(0,4.0),
     main="Correlation after PCA",xlab="Spearman Correlation")


```
### EFA

```{r  results = "asis", warning = FALSE, dpi=600, fig.height= 5, fig.width= 6}
par(mfrow=c(1,1),cex=0.6,cex.main=0.8)

## Train Correlation

cormat <- cor(EFA_Train,method="pearson")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

pander::pander(c(Train=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(EFA_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

## Test Correlation
cormat <- cor(EFA_Predicted,method="spearman")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

pander::pander(c(Test=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(EFA_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

gplots::heatmap.2(abs(cormat),
                  trace = "none",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  breaks = breaks,
                  main = "Correlation after EFA",
                  cexRow = 0.25,
                  cexCol = 0.25,
                  srtCol=35,
                  srtRow=75,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")


hist(cormat,freq=FALSE,
     density=NULL,
     xlim=c(-1,1),
     ylim=c(0,4.0),
     main="Correlation after EFA",xlab="Spearman Correlation")

```

### PCA Whitening

```{r results = "asis", warning = FALSE, dpi=600, fig.height= 5, fig.width= 6}


## Train Correlation

cormat <- cor(PCAWhite_Train,method="pearson")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

pander::pander(c(Train=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(PCAWhite_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

## Test Correlation
cormat <- cor(PCAWhitePredicted,method="spearman")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

pander::pander(c(Test=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(PCAWhite_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

gplots::heatmap.2(abs(cormat),
                  trace = "none",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  breaks = breaks,
                  main = "Correlation after PCAWhite",
                  cexRow = 0.25,
                  cexCol = 0.25,
                  srtCol=35,
                  srtRow=75,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")


hist(cormat,freq=FALSE,
     density=NULL,
     xlim=c(-1,1),
     ylim=c(0,4.0),
     main="Correlation after PCAWhite",xlab="Spearman Correlation")

```


### The tables

```{r results = "asis"}

pander::pander(univar$orderframe[1:TopVariables,univariate_columns])
pander::pander(univarDe$orderframe[1:TopVariables,univariate_columns])
pander::pander(univarDeDri$orderframe[1:TopVariables,univariate_columns])
pander::pander(univarDeDriSpear$orderframe[1:TopVariables,univariate_columns])
pander::pander(univarPCA$orderframe[1:TopVariables,univariate_columns])
pander::pander(univarEFA$orderframe[1:TopVariables,univariate_columns])
pander::pander(univarWhite$orderframe[1:TopVariables,univariate_columns])

topUni <- univar$orderframe$Name[1:TopVariables]
topDe <- univarDe$orderframe$Name[1:TopVariables]
topDeDri <- univarDeDri$orderframe$Name[1:TopVariables]
topDeDriSpear <- univarDeDriSpear$orderframe$Name[1:TopVariables]
topPCA <- univarPCA$orderframe$Name[1:TopVariables]
topEFA <- univarEFA$orderframe$Name[1:TopVariables]
topPCAWhite <- univarWhite$orderframe$Name[1:TopVariables]
```

### Model of top variables

```{r results = "asis", warning = FALSE, dpi=600, fig.height= 4, fig.width= 8}
par(mfrow=c(1,2),cex=0.6)

lmRAW <- glm(paste(outcome,"~."),
             trainDataFrame[,c(outcome,topUni)],
             family="binomial")
pr <- predictionStats_binary(cbind(testDataFrame[,outcome],predict(lmRAW,testDataFrame)),"Top Raw",cex=0.75)

lmDe <- glm(paste(outcome,"~."),
            DEdataframe[,c(outcome,topDe)],
            family="binomial")
pr <- predictionStats_binary(cbind(predTestDe[,outcome],predict(lmDe,predTestDe)),"Top Blind:GDSTM",cex=0.75)

lmDri <- glm(paste(outcome,"~."),
            DriDEdataframe[,c(outcome,topDeDri)],
            family="binomial")
pr <- predictionStats_binary(cbind(predTestDe[,outcome],predict(lmDri,predTestDri)),"Top Driven:GDSTM",cex=0.75)

lmDriSpear <- glm(paste(outcome,"~."),
            DriDEdataframeSpear[,c(outcome,topDeDriSpear)],
            family="binomial")
pr <- predictionStats_binary(cbind(predTestDriSpear[,outcome],predict(lmDriSpear,predTestDriSpear)),"Spearman Top Driven:GDSTM",cex=0.7)


lmPCA <- glm(paste(outcome,"~."),
            PCA_Train[,c(outcome,topPCA)],
            family="binomial")
pr <- predictionStats_binary(cbind(PCA_Predicted[,outcome],predict(lmPCA,PCA_Predicted)),"Top PCA",cex=0.75)


lmEFA <- glm(paste(outcome,"~."),
            EFA_Train[,c(outcome,topEFA)],
            family="binomial")
pr <- predictionStats_binary(cbind(EFA_Predicted[,outcome],predict(lmEFA,EFA_Predicted)),"Top EFA",cex=0.75)


lmPCAW <- glm(paste(outcome,"~."),
            PCAWhite_Train[,c(outcome,topPCAWhite)],
            family="binomial")
pr <- predictionStats_binary(cbind(PCAWhitePredicted[,outcome],predict(lmPCAW,PCAWhitePredicted)),"Top PCA_Cor",cex=0.75)

```

