---
title: "IDeA, EFA: Cross Validation Parkinson's"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
  word_document: 
    reference_docx: WordStyle_FRESA.docx
    toc: yes
    fig_caption: yes
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message = FALSE,comment = "#>")
```

# Parkinson's EFA-IDeA

### Loading the libraries

```{r}
library("FRESA.CAD")
library(psych)
library(whitening)
library("vioplot")

library(readxl)
op <- par(no.readonly = TRUE)
pander::panderOptions('digits', 3)
pander::panderOptions('table.split.table', 400)
pander::panderOptions('keep.trailing.zeros',TRUE)

```


## The Data

```{r}

pd_speech_features <- as.data.frame(read_excel("~/GitHub/FCA/Data/pd_speech_features.xlsx",sheet = "pd_speech_features", range = "A2:ACB758"))




```

### The Average of the Three Repetitions

Each subject had three repeated observations. Here I'll use the average of the three experiments per subject.

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 6.0, fig.width= 8.0}
rep1Parkison <- subset(pd_speech_features,RID==1)
rownames(rep1Parkison) <- rep1Parkison$id
rep1Parkison$id <- NULL
rep1Parkison$RID <- NULL
rep1Parkison[,1:ncol(rep1Parkison)] <- sapply(rep1Parkison,as.numeric)

rep2Parkison <- subset(pd_speech_features,RID==2)
rownames(rep2Parkison) <- rep2Parkison$id
rep2Parkison$id <- NULL
rep2Parkison$RID <- NULL
rep2Parkison[,1:ncol(rep2Parkison)] <- sapply(rep2Parkison,as.numeric)

rep3Parkison <- subset(pd_speech_features,RID==3)
rownames(rep3Parkison) <- rep3Parkison$id
rep3Parkison$id <- NULL
rep3Parkison$RID <- NULL
rep3Parkison[,1:ncol(rep3Parkison)] <- sapply(rep3Parkison,as.numeric)

whof <- !(colnames(rep1Parkison) %in% c("gender","class"));
avgParkison <- rep1Parkison;
avgParkison[,whof] <- (rep1Parkison[,whof] + rep2Parkison[,whof] + rep3Parkison[,whof])/3

pander::pander(table(avgParkison$class))


signedlog <- function(x) { return (sign(x)*log(abs(1.0e12*x)+1.0))}
whof <- !(colnames(avgParkison) %in% c("gender","class"));
avgParkison[,whof] <- signedlog(avgParkison[,whof])


```

#### Training and testing sets

```{r results = "asis"}
dataframe <- avgParkison
outcome <- "class"


outcomeIDeA <- outcome
#outcomeIDeA <- NULL
cormethod <- "fast"
trainFraction <- 0.5
cvloops <- 50
rhoThreshold <- 0.8
nEFAFactors <- 10
topFeat <- 20


set.seed(10)
colors=c("red","blue","darkblue","green","darkgreen")

binOutcome <- dataframe[,outcome]
getAUC <- function(x)
{
  return (pROC::roc(binOutcome,x,plot=FALSE,auc=TRUE,quiet = TRUE)$auc[1])
}

```


### Data specs

```{r results = "asis"}
pander::pander(c(rows=nrow(dataframe),col=ncol(dataframe)-1))
pander::pander(table(dataframe[,outcome]))


```


## Shrinkage Analysis

```{r results = "asis", warning = FALSE}

RawAUCs <- numeric()
IDeAAUCs <- numeric()
IDeAAUCs_LV <- numeric()
EFAAUCs <- numeric()
EFAAUCs_LV <- numeric()


ChangeRawAUCs <- numeric()
ChangeIDeAAUCs <- numeric()
ChangeIDeAAUCs_LV <- numeric()
ChangeEFAAUCs <- numeric()
ChangeEFAAUCs_LV <- numeric()


overlapRaw <- numeric()
overlapIDeA <- numeric()
overlapIDeA_LV <- numeric()
overlapEFA <- numeric()
overlapEFA_LV <- numeric()

cvloop <- 1
rawthecolnames <- colnames(dataframe)
varlist <- rawthecolnames[rawthecolnames != outcome]
selvarRAW <- numeric();
selvarIDeA <- numeric();
selvarEFA <- numeric();
selvarIDeA_LV <- numeric();
selvarEFA_LV <- numeric();
corvar <- numeric();
IDeALVsize <- numeric();

for (cvloop in c(1:cvloops))
{
    print(cvloop)
    foldSample <- sample(nrow(dataframe),nrow(dataframe)*trainFraction)
    
    foldOne <- dataframe[foldSample,]
    foldTwo <- dataframe[-foldSample,]
#    pander::pander(table(foldOne[,outcome]))
#    pander::pander(table(foldTwo[,outcome]))
    
    #################### Fold One ##################################
    topSigfeatOnep <- univariate_Wilcoxon(foldOne,outcome,limit=0,thr=rhoThreshold)
    selvarRAW <- c(selvarRAW,length(topSigfeatOnep))
    topSigfeatOne <- names(topSigfeatOnep)
    if (length(topSigfeatOne)>topFeat)
    {
      topSigfeatOne <- topSigfeatOne[1:topFeat]
    }
    

    #################### Fold Two ##################################
    topSigfeatTwo <- names(univariate_Wilcoxon(foldTwo,outcome,limit=-1,thr=0.999))

    ##################### Overlap #################################
    
    botFoldFeat <- unique(c(topSigfeatOne,topSigfeatTwo))
    
    binOutcome <- foldOne[,outcome]
    foldAUCOne <- apply(as.data.frame(foldOne[,botFoldFeat]),2,getAUC)
    names(foldAUCOne) <- botFoldFeat

    binOutcome <- foldTwo[,outcome]
    foldAUCTwo <- apply(as.data.frame(foldTwo[,botFoldFeat]),2,getAUC)
    names(foldAUCTwo) <- botFoldFeat

    
    ChangeRawAUCs <- c(ChangeRawAUCs,
                       foldAUCTwo[topSigfeatOne] - foldAUCOne[topSigfeatOne]
                      )

    RawAUCs <- c(RawAUCs,foldAUCTwo[topSigfeatOne])

    topSigfeatOne <- names(topSigfeatOnep)
    if (topSigfeatOnep[1]<0.2)
    {
      overlapRaw <- c(overlapRaw,sum(topSigfeatOne %in% topSigfeatTwo)/length(topSigfeatOne))
    }
    else
    {
       overlapRaw <- c(overlapRaw,0)
    }

    #################################### IDeA ################################
    
    IDeAData <- IDeA(foldOne,thr=rhoThreshold,method=cormethod,Outcome=outcomeIDeA)
    predIDeAData <- predictDecorrelate(IDeAData,foldTwo)
    
    
    topSigfeatOnep <- univariate_Wilcoxon(IDeAData,outcome,limit=0,thr=rhoThreshold)
    selvarIDeA <- c(selvarIDeA,length(topSigfeatOnep))

    topSigfeatOne <- names(topSigfeatOnep)
    topSigfeatTwo <- names(univariate_Wilcoxon(predIDeAData,outcome,limit=-1,thr=0.999))
    botFoldFeat <- unique(c(topSigfeatOne,topSigfeatTwo))

    
    LVars <- getLatentCoefficients(IDeAData);
    IDeALVsize <- c(IDeALVsize,mean(sapply(LVars,length)))
    
    binOutcome <- foldOne[,outcome]
    foldAUCOne <- apply(as.data.frame(IDeAData[,botFoldFeat]),2,getAUC)
    names(foldAUCOne) <- botFoldFeat

    binOutcome <- foldTwo[,outcome]
    foldAUCTwo <- apply(as.data.frame(predIDeAData[,botFoldFeat]),2,getAUC)
    names(foldAUCTwo) <- botFoldFeat

    ### Only Top AUC
    topOneFeat <- topSigfeatOne
    if (length(topOneFeat)>topFeat)
    {
      topOneFeat <- topOneFeat[1:topFeat]
    }
    
    ChangeIDeAAUCs <- c(ChangeIDeAAUCs,
                        foldAUCTwo[topOneFeat] - foldAUCOne[topOneFeat]
                      )

    IDeAAUCs <- c(IDeAAUCs,foldAUCTwo[topOneFeat])

    
    #################### Only Latent Analysis #########################
    latentVar <- colnames(IDeAData)[!(colnames(IDeAData) %in% rawthecolnames)]
    selvarIDeA_LV <- c(selvarIDeA_LV,sum(topSigfeatOne %in% latentVar))
    
    top_LV <-topOneFeat[topOneFeat %in% latentVar]
    if (length(top_LV) > 0)
    {
      ChangeIDeAAUCs_LV <- c(ChangeIDeAAUCs_LV,
                         foldAUCTwo[top_LV] - foldAUCOne[top_LV]
                        )
  
      IDeAAUCs_LV <- c(IDeAAUCs_LV,foldAUCTwo[top_LV])
    }
    
    if (topSigfeatOnep[1]<0.2)
    {
      overlapIDeA <- c(overlapIDeA,sum(topSigfeatOne %in% topSigfeatTwo)/length(topSigfeatOne))
    }
    else
    {
      overlapIDeA <- c(overlapIDeA,0)
    }

    sigOne_LV <- topSigfeatOne[topSigfeatOne %in% latentVar]
    sigTwo_LV <- topSigfeatTwo[topSigfeatTwo %in% latentVar]
    if ((topSigfeatOnep[1]<0.2) && (length(sigOne_LV)>0))
    {
      overlapIDeA_LV <- c(overlapIDeA_LV,sum(sigOne_LV %in% sigTwo_LV)/length(sigOne_LV))
    }
    else
    {
      overlapIDeA_LV <- c(overlapIDeA_LV,0)
    }
    

    ####################### EFA Viable Features #############################
    
    iscontinous <- sapply(apply(foldOne,2,unique),length) > 5 ## Only variables with enough samples
    
    noclassData <- foldOne[,iscontinous]
    cmat <- cor(noclassData);
    diag(cmat) <- 0;
    maxcor <- apply(cmat,2,max);
    sumcor <- apply(cmat>=rhoThreshold,2,sum)
    topcor <- names(maxcor[sumcor > (1 + 0.5*nEFAFactors)]) ## Only correlated features will create latent variables
    cmat <- NULL
    corvar <- c(corvar,length(topcor))
    notcorr <- colnames(foldOne)[!(colnames(foldOne) %in% topcor)]
    noclassData <- noclassData[,topcor]
    noclassDataTest <- foldTwo[,topcor]
        
    ## EFA Fold One to Two
    
    uls <- try(fa(noclassData,nEFAFactors,rotate="varimax"))  #unweighted least squares is minres 
    if (!inherits(uls, "try-error"))
    {
      EFA_Train <- as.data.frame(cbind(predict(uls,noclassData),foldOne[,notcorr]))
      colnames(EFA_Train) <- c(colnames(predict(uls,noclassData)),notcorr)
      EFA_Predicted <- as.data.frame(cbind(predict(uls,noclassDataTest),foldTwo[,notcorr]))
      colnames(EFA_Predicted) <- c(colnames(predict(uls,noclassDataTest)),notcorr)
      
  
      topSigfeatOnep <- univariate_Wilcoxon(EFA_Train,outcome,limit=0,thr=rhoThreshold)
      topSigfeatOne <- names(topSigfeatOnep)
      topSigfeatTwo <- names(univariate_Wilcoxon(EFA_Predicted,outcome,limit=-1,thr=0.999))
      selvarEFA <- c(selvarEFA,length(topSigfeatOnep))
  

      botFoldFeat <- unique(c(topSigfeatOne,topSigfeatTwo))
      
      binOutcome <- foldOne[,outcome]
      foldAUCOne <- apply(as.data.frame(EFA_Train[,botFoldFeat]),2,getAUC)
      names(foldAUCOne) <- botFoldFeat
  
      binOutcome <- foldTwo[,outcome]
      foldAUCTwo <- apply(as.data.frame(EFA_Predicted[,botFoldFeat]),2,getAUC)
      names(foldAUCTwo) <- botFoldFeat
      
      topOneF <- topSigfeatOne
  
      if (length(topOneF)>topFeat)
      {
        topOneF <- topOneF[1:topFeat]
      }

      ChangeEFAAUCs <- c(ChangeEFAAUCs,
                         foldAUCTwo[topOneF]- foldAUCOne[topOneF]
                        )
  
      EFAAUCs <- c(EFAAUCs,foldAUCTwo[topOneF])
      
      if (topSigfeatOnep[1]<0.2)
      {
        overlapEFA <- c(overlapEFA,sum(topSigfeatOne %in% topSigfeatTwo)/length(topSigfeatOne))
      }
      else
      {
        overlapEFA <- c(overlapEFA,0)
      }
      
      #################### Only Latent Analysis #########################
      latentVar <- colnames(EFA_Train)[!(colnames(EFA_Train) %in% rawthecolnames)]
      selvarEFA_LV <- c(selvarEFA_LV,sum(topSigfeatOne %in% latentVar))
  
      topSigfeatOne <- topSigfeatOne[topSigfeatOne %in% latentVar]
      topSigfeatTwo <- topSigfeatTwo[topSigfeatTwo %in% latentVar]
      
      topOneLat <- topOneF[topOneF %in% latentVar]
  
      if (length(topOneLat)>0)
      {
        ChangeEFAAUCs_LV <- c(ChangeEFAAUCs_LV,
                           foldAUCTwo[topOneLat] - foldAUCOne[topOneLat]
                          )
    
        EFAAUCs_LV <- c(EFAAUCs_LV,foldAUCTwo[topOneLat])
      }
      
      if ((topSigfeatOnep[1]<0.2) && (length(topSigfeatOne) > 0))
      {
        overlapEFA_LV <- c(overlapEFA_LV,sum(topSigfeatOne %in% topSigfeatTwo)/length(topSigfeatOne))
      }
      else
      {
        overlapEFA_LV <- c(overlapEFA_LV,0)
      }
    }
      

    
}

RawAUCs <- RawAUCs[!is.na(RawAUCs)]
IDeAAUCs <- IDeAAUCs[!is.na(IDeAAUCs)]
IDeAAUCs_LV <- IDeAAUCs_LV[!is.na(IDeAAUCs_LV)]
EFAAUCs <- EFAAUCs[!is.na(EFAAUCs)]
EFAAUCs_LV <- EFAAUCs_LV[!is.na(EFAAUCs_LV)]


```
### The number of discoveries

```{r results = "asis"}

pander::pander(summary(IDeALVsize),caption="Latent Variable Size")

pander::pander(summary(corvar),caption="EFA Variable Size")

pander::pander(as.matrix(c(Raw=mean(selvarRAW),
                 IDeA=mean(selvarIDeA),
                 IDeA_LV=mean(selvarIDeA_LV),
                 EFA=mean(selvarEFA),
                 EFA_LV=mean(selvarEFA_LV)
               )),caption="Average Number of Selected Variables")

```


### Violin of ROC AUC 
```{r results = "asis", warning = FALSE, dpi=300, fig.height= 4, fig.width= 5}

par(op)
par(mfrow=c(1,1),cex=0.7)

AUCs <- list(RAW=RawAUCs,IDeA=IDeAAUCs,IDeA_LV=IDeAAUCs_LV,EFA=EFAAUCs,EFA_LV=EFAAUCs_LV)
vioplot(AUCs,
        ylim=c(0.5,1.0),
        ylab="Test AUC",
        main="Test ROC AUC",
        col=colors,
        cex.axis=0.8
)
#stripchart(diffAUC, method = "jitter", col = "gray",
#           vertical = TRUE, pch = 1, add = TRUE,cex=0.35)
points(1:length(AUCs),lapply(AUCs,mean), pch = 18, col = "yellow", cex = 2.25)
par(op)


bp<- boxplot(AUCs,plot=FALSE)
bpstats <- as.data.frame(bp$stats)
rownames(bpstats) <- c("lower whisker","lower hinge","median","upper hinge","upper whisker")
colnames(bpstats) <- bp$names
pander::pander(bpstats)

```


### Violin of ROC AUC Changes
```{r results = "asis", warning = FALSE, dpi=300, fig.height= 4, fig.width= 5}

par(op)
colors=c("red","blue","darkblue","green","darkgreen")
par(mfrow=c(1,1),cex=0.7)

diffAUC <- list(RAW=ChangeRawAUCs,IDeA=ChangeIDeAAUCs,IDeA_LV=ChangeIDeAAUCs_LV,EFA=ChangeEFAAUCs,EFA_LV=ChangeEFAAUCs_LV)
vioplot(diffAUC,
        ylim=c(-0.5,0.5),
        ylab="Test-Train",
        main="Test-Train Paired ROC AUC",
        col=colors,
        cex.axis=0.8
)
#stripchart(diffAUC, method = "jitter", col = "gray",
#           vertical = TRUE, pch = 1, add = TRUE,cex=0.35)
points(1:length(diffAUC),lapply(diffAUC,mean), pch = 18, col = "yellow", cex = 2.25)
par(op)


bp<- boxplot(diffAUC,plot=FALSE)
bpstats <- as.data.frame(bp$stats)
rownames(bpstats) <- c("lower whisker","lower hinge","median","upper hinge","upper whisker")
colnames(bpstats) <- bp$names
pander::pander(bpstats)

```


### Violin of Overlap
```{r results = "asis", warning = FALSE, dpi=300, fig.height= 4, fig.width= 5}


par(op)
par(mfrow=c(1,1),cex=0.7)

posFind <- list(RAW=overlapRaw,IDeA=overlapIDeA,IDeA_LV=overlapIDeA_LV,EFA=overlapEFA,EFA_LV=overlapEFA_LV)
vioplot(posFind,
        ylim=c(0,1.0),
        ylab="Overlap",
        main="Selection Overlap",
        col=colors,
        cex.axis=0.8
)
stripchart(posFind, method = "jitter", col = "gray",
           vertical = TRUE, pch = 1, add = TRUE,cex=0.5)
points(1:length(posFind),lapply(posFind,mean), pch = 18, col = "yellow", cex = 2.25)
par(op)


bp<- boxplot(posFind,plot=FALSE)
bpstats <- as.data.frame(bp$stats)
rownames(bpstats) <- c("lower whisker","lower hinge","median","upper hinge","upper whisker")
colnames(bpstats) <- bp$names
pander::pander(bpstats)

```

### Violin of Booth Positive top 
```{r results = "asis", warning = FALSE, dpi=300, fig.height= 4, fig.width= 5}


par(op)
par(mfrow=c(1,1),cex=0.7)

posFind <- list(RAW=overlapRaw,IDeA=overlapIDeA,EFA=overlapEFA)
vioplot(posFind,
        ylim=c(0,1.0),
        ylab="Overlap",
        main="Selection Overlap",
        col=colors[c(1,2,4)],
        cex.axis=0.8
)
stripchart(posFind, method = "jitter", col = "gray",
           vertical = TRUE, pch = 1, add = TRUE,cex=0.5)
points(1:length(posFind),lapply(posFind,mean), pch = 18, col = "yellow", cex = 2.25)
par(op)


bp<- boxplot(posFind,plot=FALSE)
bpstats <- as.data.frame(bp$stats)
rownames(bpstats) <- c("lower whisker","lower hinge","median","upper hinge","upper whisker")
colnames(bpstats) <- bp$names
pander::pander(bpstats)

```
