---
title: "PCA EFA: DARWIN"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
  word_document: 
    reference_docx: WordStyle_FRESA.docx
    toc: yes
    fig_caption: yes
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message = FALSE,comment = "#>")
```

# DARWIN Univariate

### Loading the libraries

```{r}
library("FRESA.CAD")
library(psych)
library(whitening)

library(readxl)
op <- par(no.readonly = TRUE)
pander::panderOptions('digits', 3)
pander::panderOptions('table.split.table', 400)
pander::panderOptions('keep.trailing.zeros',TRUE)

```



## The Data

```{r}
DARWIN <- read.csv("~/GitHub/FCA/Data/DARWIN/DARWIN.csv")
rownames(DARWIN) <- DARWIN$ID
DARWIN$ID <- NULL
DARWIN$class <- 1*(DARWIN$class=="P")
print(table(DARWIN$class))

DARWIN[,1:ncol(DARWIN)] <- sapply(DARWIN,as.numeric)

#signedlog <- function(x) { return (sign(x)*log(abs(1.0e12*x)+1.0))}
#whof <- !(colnames(DARWIN) %in% c("class"));
#DARWIN[,whof] <- signedlog(DARWIN[,whof])



```


#### Standarize the names for the reporting

```{r results = "asis"}
dataframe <- DARWIN
outcome <- "class"

trainFraction <- 0.5
rhoThreshold <- 0.5
TopVariables <- 5

set.seed(5)
trainSample <- sample(nrow(dataframe),nrow(dataframe)*trainFraction)

trainDataFrame <- dataframe[trainSample,]
testDataFrame <- dataframe[-trainSample,]

```



```{r results = "asis"}
pander::pander(c(rows=nrow(dataframe),col=ncol(dataframe)-1))
pander::pander(table(dataframe[,outcome]))

varlist <- colnames(dataframe)
varlist <- varlist[varlist != outcome]
varlist <- as.data.frame(cbind(name=varlist,desc=varlist))

```



## Univariate

```{r results = "asis"}

univariate_columns <- c("caseMean","caseStd","controlMean","controlStd","controlKSP","ROCAUC","wilcox.Zvalue")
univar <- uniRankVar(varlist,
	           paste(outcome,"~1"),
	           outcome,
	           trainDataFrame,
              rankingTest = "AUC")



```

## Decorrelation with GDSTM Blind

```{r results = "asis"}
DEdataframe <- GDSTMDecorrelation(trainDataFrame,thr=rhoThreshold)
predTestDe <- predictDecorrelate(DEdataframe,testDataFrame)

ltvar <- getLatentCoefficients(DEdataframe);
pander::pander(head(ltvar))
pander::pander(c(Avlen=mean(sapply(ltvar,length))))
pander::pander(c(Latent=length(ltvar)))


varlistDe <-  colnames(DEdataframe)[colnames(DEdataframe) != outcome];
univarDe <- uniRankVar(varlistDe,
              paste(outcome,"~1"),
	            outcome,
              DEdataframe,
              rankingTest = "AUC")


```

## Decorrelation with GDSTM Blind/Spearman

```{r results = "asis"}
DEdataframeSpear <- GDSTMDecorrelation(trainDataFrame,thr=rhoThreshold,method="spearman")
predTestDeSpear <- predictDecorrelate(DEdataframeSpear,testDataFrame)

ltvar <- getLatentCoefficients(DEdataframeSpear);
pander::pander(head(ltvar))
pander::pander(c(Avlen=mean(sapply(ltvar,length))))
pander::pander(c(Latent=length(ltvar)))


varlistDeSpear <-  colnames(DEdataframeSpear)[colnames(DEdataframeSpear) != outcome];
univarDeSpear <- uniRankVar(varlistDeSpear,
              paste(outcome,"~1"),
	            outcome,
              DEdataframeSpear,
              rankingTest = "AUC")


```

## Decorrelation with GDSTM Driven

```{r results = "asis"}

DriDEdataframe <- GDSTMDecorrelation(trainDataFrame,Outcome=outcome,thr=rhoThreshold)
predTestDri <- predictDecorrelate(DriDEdataframe,testDataFrame)


ltvar <- getLatentCoefficients(DriDEdataframe);
pander::pander(head(ltvar))
pander::pander(c(Avlen=mean(sapply(ltvar,length))))
pander::pander(c(Latent=length(ltvar)))


varlistDe <-  colnames(DriDEdataframe)[colnames(DriDEdataframe) != outcome];
univarDeDri <- uniRankVar(varlistDe,
              paste(outcome,"~1"),
	            outcome,
              DriDEdataframe,
              rankingTest = "AUC")


```

## Decorrelation with GDSTM Driven and Spearman

```{r results = "asis"}

DriDEdataframeSpear <- GDSTMDecorrelation(trainDataFrame,Outcome=outcome,thr=rhoThreshold,method="spearman")
predTestDriSpear <- predictDecorrelate(DriDEdataframeSpear,testDataFrame)


ltvar <- getLatentCoefficients(DriDEdataframeSpear);
pander::pander(head(ltvar))
pander::pander(c(Avlen=mean(sapply(ltvar,length))))
pander::pander(c(Latent=length(ltvar)))


varlistDeSpear <-  colnames(DriDEdataframeSpear)[colnames(DriDEdataframeSpear) != outcome];
univarDeDriSpear <- uniRankVar(varlistDeSpear,
              paste(outcome,"~1"),
	            outcome,
              DriDEdataframeSpear,
              rankingTest = "AUC")

```

### Get continous correlated features
```{r}
iscontinous <- sapply(apply(trainDataFrame,2,unique),length) > 5 ## Only variables with enough samples

noclassData <- trainDataFrame[,iscontinous]
cmat <- cor(noclassData);
diag(cmat) <- 0;
maxcor <- apply(cmat,2,max);
topcor <- names(maxcor[maxcor>rhoThreshold]) ## Only correlated features will be PCA
pander::pander(c(Ncor=length(topcor)))
cmat <- NULL

notcorr <- colnames(trainDataFrame)[!(colnames(trainDataFrame) %in% topcor)]
noclassData <- noclassData[,topcor]
noclassTestData <- testDataFrame[,topcor]

```

## PCA Analysis

```{r results = "asis"}

### PCA 

pc <- principal(noclassData,2*TopVariables,rotate="varimax")   #principal components
pander::pander(t(pc$loadings[1:TopVariables,1:TopVariables]))
PCA_Train <- as.data.frame(cbind(predict(pc,noclassData),trainDataFrame[,notcorr]))
colnames(PCA_Train) <- c(colnames(predict(pc,noclassData)),notcorr)

PCA_Predicted <- as.data.frame(cbind(predict(pc,noclassTestData),testDataFrame[,notcorr]))
colnames(PCA_Predicted) <- c(colnames(predict(pc,noclassTestData)),notcorr)

varlistPCA <-  colnames(PCA_Predicted)[colnames(PCA_Predicted) != "class"];
univarPCA <- uniRankVar(varlistPCA,
              paste(outcome,"~1"),
	            outcome,
              PCA_Train,
              rankingTest = "AUC")



```

## EFA
```{r results = "asis"}

uls <- fa(noclassData,2*TopVariables,rotate="varimax")  #unweighted least squares is minres 
pander::pander(t(uls$weights[1:TopVariables,1:TopVariables])) 
EFA_Train <- as.data.frame(cbind(predict(uls,noclassData),trainDataFrame[,notcorr]))
colnames(EFA_Train) <- c(colnames(predict(uls,noclassData)),notcorr)
EFA_Predicted <- as.data.frame(cbind(predict(uls,noclassTestData),testDataFrame[,notcorr]))
colnames(EFA_Predicted) <- c(colnames(predict(uls,noclassTestData)),notcorr)

varlistEFA <-  colnames(EFA_Predicted)[colnames(EFA_Predicted) != "class"];
univarEFA <- uniRankVar(varlistEFA,
              paste(outcome,"~1"),
	            outcome,
              EFA_Train,
              rankingTest = "AUC")


```


## White
```{r results = "asis"}
WhiteMat = whiteningMatrix(cov(noclassData), method="PCA")
tokeep <- apply(is.na(WhiteMat),1,sum) == 0
WhiteMat <- WhiteMat[tokeep,]
sum(is.na(WhiteMat))
tokeep <- apply(abs(WhiteMat),1,sum) < 1.0e6
WhiteMat <- WhiteMat[tokeep,]
#apply(abs(WhiteMat),1,sum)

pander::pander(c(ncol=ncol(WhiteMat),nrow=nrow(WhiteMat)))

pander::pander(WhiteMat[1:TopVariables,1:TopVariables]) 
PCAWhite_Train <- as.data.frame(cbind(tcrossprod(as.matrix(noclassData), WhiteMat),trainDataFrame[,notcorr]))
colnames(PCAWhite_Train) <- c(colnames(tcrossprod(as.matrix(noclassData), WhiteMat)),notcorr)

PCAWhitePredicted <- as.data.frame(cbind(tcrossprod(as.matrix(noclassTestData), WhiteMat),testDataFrame[,notcorr]))
colnames(PCAWhitePredicted) <- c(colnames(tcrossprod(as.matrix(noclassTestData), WhiteMat)),notcorr)

varlistWhite <-  colnames(PCAWhitePredicted)[colnames(PCAWhitePredicted) != "class"];
univarWhite <- uniRankVar(varlistWhite,
              paste(outcome,"~1"),
	            outcome,
              PCAWhite_Train,
              rankingTest = "AUC")


```

## Correlation Matrices

### RAW

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5, fig.width= 6}
par(cex=1.0,cex.main=0.8)
breaks <- c(0:5)/5.0;

cormat <- cor(testDataFrame,method="spearman")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;
pander::pander(max(abs(cormat)))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(Raw_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

gplots::heatmap.2(abs(cormat),
                  trace = "none",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  breaks = breaks,
                  main = "Raw Correlation",
                  cexRow = 0.25,
                  cexCol = 0.25,
                  srtCol=35,
                  srtRow=75,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature"
                  )

#hist(cormat,freq=FALSE,
#     density=NULL,
#     xlim=c(-1,1),
#     ylim=c(0,4.0),
#     main="Raw Correlation",xlab="Spearman Correlation")

rawDen <- density(cormat,from=-1,to=1)
par(op)
```

### GDSTM Blind

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5, fig.width= 6}
par(cex=1.0,cex.main=0.8)

## Train Correlation

cormat <- cor(DEdataframe,method="pearson")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

#pander::pander(colnames(cormat)[apply(abs(cormat),2,max)>rhoThreshold])

pander::pander(c(Train=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(HCA_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

## Test Correlation
cormat <- cor(predTestDe,method="spearman")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

pander::pander(c(Test=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(HCA_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

gplots::heatmap.2(abs(cormat),
                  trace = "none",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  breaks = breaks,
                  main = "Correlation after HCA",
                  cexRow = 0.25,
                  cexCol = 0.25,
                  srtCol=35,
                  srtRow=75,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")

#hist(cormat,freq=FALSE,
#     density=NULL,
#     xlim=c(-1,1),
#     ylim=c(0,4.0),
#     main="Correlation after GDSTM",xlab="Spearman Correlation")

DeDen <- density(cormat,from=-1,to=1)


par(op)

```

### GDSTM Blind/Spearman

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5, fig.width= 6}
par(cex=1.0,cex.main=0.8)

## Train Correlation

cormat <- cor(DEdataframeSpear,method="pearson")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

#pander::pander(colnames(cormat)[apply(abs(cormat),2,max)>rhoThreshold])

pander::pander(c(Train=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(HCA_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

## Test Correlation
cormat <- cor(predTestDeSpear,method="spearman")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

pander::pander(c(Test=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(HCA_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

gplots::heatmap.2(abs(cormat),
                  trace = "none",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  breaks = breaks,
                  main = "Correlation after HCA:Spearman",
                  cexRow = 0.25,
                  cexCol = 0.25,
                  srtCol=35,
                  srtRow=75,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")

#hist(cormat,freq=FALSE,
#     density=NULL,
#     xlim=c(-1,1),
#     ylim=c(0,4.0),
#     main="Correlation after GDSTM",xlab="Spearman Correlation")

DeSpearDen <- density(cormat,from=-1,to=1)

par(op)


```

### GDSTM Driven

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5, fig.width= 6}
par(cex=1.0,cex.main=0.8)

## Train Correlation

cormat <- cor(DriDEdataframe,method="pearson")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;
#pander::pander(colnames(cormat)[apply(abs(cormat),2,max)>rhoThreshold])

pander::pander(c(Train=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(HCA_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

## Test Correlation
cormat <- cor(DriDEdataframe,method="spearman")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

pander::pander(c(Test=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(HCA_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

gplots::heatmap.2(abs(cormat),
                  trace = "none",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  breaks = breaks,
                  main = "Correlation after Driven-HCA",
                  cexRow = 0.25,
                  cexCol = 0.25,
                  srtCol=35,
                  srtRow=75,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")
#hist(cormat,freq=FALSE,
#     density=NULL,
#     xlim=c(-1,1),
#     ylim=c(0,4.0),
#     main="Correlation after Driven-GDSTM",xlab="Spearman Correlation")

DeDrivDen <- density(cormat,from=-1,to=1)
par(op)


```

### GDSTM Spearman
```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5, fig.width= 6}
par(cex=1.0,cex.main=0.8)

## Train Correlation

cormat <- cor(DriDEdataframeSpear,method="pearson")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;
#pander::pander(colnames(cormat)[apply(abs(cormat),2,max)>rhoThreshold])

pander::pander(c(Train=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(HCAS_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

## Test Correlation

cormat <- cor(predTestDriSpear,method="spearman")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

pander::pander(c(Test=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(HCAS_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

gplots::heatmap.2(abs(cormat),
                  trace = "none",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  breaks = breaks,
                  main = "Correlation: Driven/Spearman",
                  cexRow = 0.25,
                  cexCol = 0.25,
                  srtCol=35,
                  srtRow=75,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")

#hist(cormat,freq=FALSE,
#     density=NULL,
#     xlim=c(-1,1),
#     ylim=c(0,4.0),
#     main="Correlation after GDSTM with Spearman",xlab="Spearman Correlation")

DeDrivSpearDen <- density(cormat,from=-1,to=1)
par(op)

```

### PCA

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5, fig.width= 6}
par(cex=1.0,cex.main=0.8)



## Train Correlation

cormat <- cor(PCA_Train,method="pearson")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

pander::pander(c(Train=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(PCA_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

## Test Correlation
cormat <- cor(PCA_Predicted,method="spearman")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

pander::pander(c(Test=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(PCA_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

gplots::heatmap.2(abs(cormat),
                  trace = "none",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  breaks = breaks,
                  main = "Correlation after PCA",
                  cexRow = 0.25,
                  cexCol = 0.25,
                  srtCol=35,
                  srtRow=75,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")


#hist(cormat,freq=FALSE,
#     density=NULL,
#     xlim=c(-1,1),
#     ylim=c(0,4.0),
#     main="Correlation after PCA",xlab="Spearman Correlation")

PCADen <- density(cormat,from=-1,to=1)

par(op)

```
### EFA

```{r  results = "asis", warning = FALSE, dpi=300, fig.height= 5, fig.width= 6}
par(cex=1.0,cex.main=0.8)

## Train Correlation

cormat <- cor(EFA_Train,method="pearson")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

pander::pander(c(Train=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(EFA_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

## Test Correlation
cormat <- cor(EFA_Predicted,method="spearman")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

pander::pander(c(Test=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(EFA_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

gplots::heatmap.2(abs(cormat),
                  trace = "none",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  breaks = breaks,
                  main = "Correlation after EFA",
                  cexRow = 0.25,
                  cexCol = 0.25,
                  srtCol=35,
                  srtRow=75,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")


#hist(cormat,freq=FALSE,
#     density=NULL,
#     xlim=c(-1,1),
#     ylim=c(0,4.0),
#     main="Correlation after EFA",xlab="Spearman Correlation")

EFADen <- density(cormat,from=-1,to=1)
par(op)

```

### PCA Whitening

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5, fig.width= 6}


## Train Correlation

cormat <- cor(PCAWhite_Train,method="pearson")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

pander::pander(c(Train=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(PCAWhite_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

## Test Correlation
cormat <- cor(PCAWhitePredicted,method="spearman")
cormat[is.na(cormat)] <- 0
diag(cormat) <- 0;

pander::pander(c(Test=max(abs(cormat))))
pander::pander(quantile(abs(cormat),probs=c(0.9,0.95,0.975,0.99,0.999)))
pander::pander(c(PCAWhite_fraction=sum(abs(cormat)>rhoThreshold)/ncol(cormat)/ncol(cormat)))

gplots::heatmap.2(abs(cormat),
                  trace = "none",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  breaks = breaks,
                  main = "Correlation after PCAWhite",
                  cexRow = 0.25,
                  cexCol = 0.25,
                  srtCol=35,
                  srtRow=75,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")


#hist(cormat,freq=FALSE,
#     density=NULL,
#     xlim=c(-1,1),
#     ylim=c(0,4.0),
#     main="Correlation after PCAWhite",xlab="Spearman Correlation")

WhiteDen <- density(cormat,from=-1,to=1)
par(op)

```

## The Density Plot

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 4, fig.width= 6}
par(cex=0.7)
plot(rawDen,
     xlim=c(-1,1),
     ylim=c(0,7.0),
     col="red",
     lty=1,
     lwd=4,
     main="Test: Correlation Distribution",xlab="Spearman Correlation")

lines(DeDen,col="blue",lty=2,lwd=4)
lines(DeSpearDen,col="green",lty=3,lwd=4)
lines(DeDrivDen,col="darkblue",lty=4,lwd=2)
lines(DeDrivSpearDen,col="darkgreen",lty=5,lwd=2)

lines(PCADen,col="purple",lty=6,lwd=1)
lines(EFADen,col="purple",lty=7,lwd=1)
lines(WhiteDen,col="gray",lty=8,lwd=1)

names=c("Raw","HCA:P","HCA:S","DHCA:P","DHCA:S","PCA","EFA","White:PCA")
colors=c("red","blue","green","blue","green","purple","purple","gray");
lines=c(1,2,3,4,5,6,7,8)
lwds=c(4,4,4,2,2,1,1,1)

legend("topleft",names,col=colors,lty=lines,lwd=lwds,cex=0.50)
par(op)

```


### The tables

```{r results = "asis"}

pander::pander(univar$orderframe[1:TopVariables,univariate_columns])
pander::pander(univarDe$orderframe[1:TopVariables,univariate_columns])
pander::pander(univarDeSpear$orderframe[1:TopVariables,univariate_columns])
pander::pander(univarDeDri$orderframe[1:TopVariables,univariate_columns])
pander::pander(univarDeDriSpear$orderframe[1:TopVariables,univariate_columns])
pander::pander(univarPCA$orderframe[1:TopVariables,univariate_columns])
pander::pander(univarEFA$orderframe[1:TopVariables,univariate_columns])
pander::pander(univarWhite$orderframe[1:TopVariables,univariate_columns])

topUni <- univar$orderframe$Name[1:TopVariables]
topDe <- univarDe$orderframe$Name[1:TopVariables]
topDeSpear <- univarDeSpear$orderframe$Name[1:TopVariables]
topDeDri <- univarDeDri$orderframe$Name[1:TopVariables]
topDeDriSpear <- univarDeDriSpear$orderframe$Name[1:TopVariables]
topPCA <- univarPCA$orderframe$Name[1:TopVariables]
topEFA <- univarEFA$orderframe$Name[1:TopVariables]
topPCAWhite <- univarWhite$orderframe$Name[1:TopVariables]
```

### Model of top variables

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 4, fig.width= 8}
par(mfrow=c(1,2),cex=0.6)

lmRAW <- glm(paste(outcome,"~."),
             trainDataFrame[,c(outcome,topUni)],
             family="binomial")
prRaw <- predictionStats_binary(cbind(testDataFrame[,outcome],predict(lmRAW,testDataFrame)),"Top Raw",cex=0.75)

lmDe <- glm(paste(outcome,"~."),
            DEdataframe[,c(outcome,topDe)],
            family="binomial")
prDe <- predictionStats_binary(cbind(predTestDe[,outcome],predict(lmDe,predTestDe)),"Top HCA:P",cex=0.75)

lmDeSpear <- glm(paste(outcome,"~."),
            DEdataframeSpear[,c(outcome,topDeSpear)],
            family="binomial")
prSpear <- predictionStats_binary(cbind(predTestDeSpear[,outcome],predict(lmDeSpear,predTestDeSpear)),"Top HCA:S",cex=0.75)

lmDri <- glm(paste(outcome,"~."),
            DriDEdataframe[,c(outcome,topDeDri)],
            family="binomial")
prDri <- predictionStats_binary(cbind(predTestDe[,outcome],predict(lmDri,predTestDri)),"Top DHCA:P",cex=0.75)

lmDriSpear <- glm(paste(outcome,"~."),
            DriDEdataframeSpear[,c(outcome,topDeDriSpear)],
            family="binomial")
prDriSpear <- predictionStats_binary(cbind(predTestDriSpear[,outcome],predict(lmDriSpear,predTestDriSpear)),"Top DHCA:S",cex=0.7)


lmPCA <- glm(paste(outcome,"~."),
            PCA_Train[,c(outcome,topPCA)],
            family="binomial")
prPCA <- predictionStats_binary(cbind(PCA_Predicted[,outcome],predict(lmPCA,PCA_Predicted)),"Top PCA",cex=0.75)


lmEFA <- glm(paste(outcome,"~."),
            EFA_Train[,c(outcome,topEFA)],
            family="binomial")
prEFA <- predictionStats_binary(cbind(EFA_Predicted[,outcome],predict(lmEFA,EFA_Predicted)),"Top EFA",cex=0.75)


lmPCAW <- glm(paste(outcome,"~."),
            PCAWhite_Train[,c(outcome,topPCAWhite)],
            family="binomial")
prWPCA <- predictionStats_binary(cbind(PCAWhitePredicted[,outcome],predict(lmPCAW,PCAWhitePredicted)),"Top White:PCA",cex=0.75)
par(op)

```

### The Performance Tables and Plots

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 4, fig.width= 6}

par(cex=0.6)

 aucs <- prRaw$aucs
  aucs <- rbind(aucs,prDe$aucs)
  aucs <- rbind(aucs,prSpear$aucs)
  aucs <- rbind(aucs,prDri$aucs)
  aucs <- rbind(aucs,prDriSpear$aucs)
  aucs <- rbind(aucs,prPCA$aucs)
  aucs <- rbind(aucs,prEFA$aucs)
  aucs <- rbind(aucs,prWPCA$aucs)

  
  rownames(aucs) <- c("RAW",
                        "HCA:P",
                        "HCA:S",
                        "DHCA:P",
                        "DHCA:S",
                        "PCA",
                        "EFA",
                        "WPCA"
                        )
  
  pander::pander(aucs)
  
  bpAUC <- barPlotCiError(as.matrix(aucs),
                          metricname = "ROC AUC",
                          thesets = "Test AUC",
                          themethod = rownames(aucs),
                          main = "ROC AUC",
                          offsets = c(0.5,1),
                          scoreDirection = ">",
                          ho=0.5,
                          args.legend = list(bg = "white",x="bottomleft",inset=c(0.0,0),cex=0.5),
                          col = terrain.colors(nrow(aucs))
                          )

  
 berror <- prRaw$berror
  berror <- rbind(berror,prDe$berror)
  berror <- rbind(berror,prSpear$berror)
  berror <- rbind(berror,prDri$berror)
  berror <- rbind(berror,prDriSpear$berror)
  berror <- rbind(berror,prPCA$berror)
  berror <- rbind(berror,prEFA$berror)
  berror <- rbind(berror,prWPCA$berror)


  rownames(berror) <- rownames(aucs)
  pander::pander(berror)

  bpBER <- barPlotCiError(as.matrix(berror),
                          metricname = "Balanced Error Rate",
                          thesets = "Test BER",
                          themethod = rownames(aucs),
                          main = "Balanced Error Rate",
                          offsets = c(0.5,1),
                          scoreDirection = "<",
                          ho=0.5,
                          args.legend = list(bg = "white",x="topleft",inset=c(0.0,0),cex=0.5),
                          col = terrain.colors(nrow(aucs))
                          )
  par(op)

```
