---
title: "Decorrelation-Based Feature Discovery: Thermal COVID_19"
author: "Jose Tamez"
date: "2022-10-02"
output:
  html_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
  word_document: 
    reference_docx: WordStyle_FRESA.docx
    toc: yes
    fig_caption: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message = FALSE,comment = "#>")

```

# Effect of GDSTM-Based Decorrelation on Feature Discovery

Here I showcase of to use BSWiMS feature selection/modeling function coupled with Goal Driven Sparse Transformation Matrix (GDSTM) as a pre-processing step to decorrelate highly correlated features. The aim(s) are:

1.  To improve model performance by uncovering the hidden information between correlated features.

2.  To simplify the interpretation of the machine learning models.

This demo will use:

-   *FRESA.CAD::GDSTMDecorrelation()*. For Decorrelation of Multidimensional data sets

    -   *FRESA.CAD::getDerivedCoefficients()*. For the extraction of the model of the newly discovered of decorrelated features.

-   *FRESA.CAD::randomCV()* For the cross-validation of the Machine Learning models

-   *FRESA.CAD::BSWiMS.model()*. For the generation of bootstrapped logistic models

    -   *FRESA.CAD::summary()*. For the summary description of the BSWiMS model

-   *FRESA.CAD::predictionStats_binary()*. For describing the performance of the model

-   *heatmap.2()*. For displaying the correlation matrix

-   *igraph::graph_from_adjacency_matrix()*. For the display of the network of BSWiMS formulas


### Loading the libraries

```{r}
library("FRESA.CAD")
library(readxl)
library(igraph)

op <- par(no.readonly = TRUE)
pander::panderOptions('digits', 3)
pander::panderOptions('table.split.table', 400)
pander::panderOptions('keep.trailing.zeros',TRUE)

```

## Material and Methods

### Signed Log Transform

The function will be used to transform all the continuous features of the data

```{r}
signedlog <- function(x) { return (sign(x)*log(abs(x)+1.0e-12))}

```

## Data: The COVID_19 Data-Set

The data to process is described in:

<https://zenodo.org/record/4156647#.Y1bSF3bMKUk>


Thermal Saliva Testing Dataset


10.5281/zenodo.4156647
<https://doi.org/10.5281/zenodo.4156647>


I added a column to the data identifying the repeated experiments.

```{r}

SalivaThermal <- as.data.frame(read_excel("~/GitHub/FCA/Data/SalivaThermal_Source_Data_2.xlsx"))


SalivaThermal_set1 <- subset(SalivaThermal,RepID==1)
rownames(SalivaThermal_set1) <- SalivaThermal_set1$ID
SalivaThermal_set1$RepID <- NULL
SalivaThermal_set1$ID <- NULL
SalivaThermal_set1$Ct <- NULL

SalivaThermal_set2 <- subset(SalivaThermal,RepID==2)
rownames(SalivaThermal_set2) <- SalivaThermal_set2$ID
SalivaThermal_set2$RepID <- NULL
SalivaThermal_set2$ID <- NULL
SalivaThermal_set2$Ct <- NULL

SalivaThermal_set3 <- subset(SalivaThermal,RepID==3)
rownames(SalivaThermal_set3) <- SalivaThermal_set3$ID
SalivaThermal_set3$RepID <- NULL
SalivaThermal_set3$ID <- NULL
SalivaThermal_set3$Ct <- NULL

SalivaThermal_Avg <- (SalivaThermal_set1 + SalivaThermal_set2 + SalivaThermal_set3)/3

SalivaThermal_d1 <- SalivaThermal_Avg[,-1] - SalivaThermal_Avg[,c(1:ncol(SalivaThermal_Avg)-1)]
#SalivaThermal_d1[,ncol(SalivaThermal_d1)] <- 0
#plot(apply(SalivaThermal_d1,2,mean))
SalivaThermal_d2 <-  -SalivaThermal_d1[,c(1:ncol(SalivaThermal_d1)-1)] + SalivaThermal_d1[,-1]
#SalivaThermal_d2[,ncol(SalivaThermal_d2)] <- 0
#plot(apply(SalivaThermal_d2,2,mean))
colnames(SalivaThermal_d2) <- paste("DD",colnames(SalivaThermal_d2),sep="_")
colnames(SalivaThermal_d1) <- paste("D",colnames(SalivaThermal_d1),sep="_")

colnames(SalivaThermal_Avg) <- paste("V",colnames(SalivaThermal_Avg),sep="_")

#SalivaThermal_Avg <- cbind(SalivaThermal_Avg,SalivaThermal_d1,SalivaThermal_d2)
#SalivaThermal_Avg <- cbind(SalivaThermal_Avg,SalivaThermal_d1,SalivaThermal_d2)
#SalivaThermal_Avg <- SalivaThermal_d2
SalivaThermal_Avg$class <- 1*(str_detect(rownames(SalivaThermal_Avg),"P"))
##The fraction of samples in the training set
pander::pander(table(SalivaThermal_Avg$class))
trainFraction=0.90

SalivaThermal_d2$class <- 1*(str_detect(rownames(SalivaThermal_d2),"P"))

```


#### Correlation Matrix of the Decorrelated Test Data

The heat map of the testing set.

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}

par(cex=0.6,cex.main=0.85,cex.axis=0.7)
cormat <- cor(SalivaThermal_Avg[,colnames(SalivaThermal_Avg) != "class"],method="spearman")
cormat[is.na(cormat)] <- 0
gplots::heatmap.2(abs(cormat),
                  trace = "none",
#                  scale = "row",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  main = "COVID-19: Infrared Signal Correlation",
                  cexRow = 0.35,
                  cexCol = 0.35,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")

## The decorrelation
deThermal <- GDSTMDecorrelation(SalivaThermal_Avg,thr=0.80,verbose = TRUE)
#deThermal <- GDSTMDecorrelation(SalivaThermal_Avg,thr=0.80,relaxed=FALSE,verbose = TRUE)

cormat <- cor(deThermal[,colnames(deThermal) != "class"],method="spearman")
cormat[is.na(cormat)] <- 0
gplots::heatmap.2(abs(cormat),
                  trace = "none",
#                  scale = "none",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  main = "COVID-19 Correlation after GDSTM",
                  cexRow = 0.35,
                  cexCol = 0.35,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")

par(op)
```

### The decorrelation matrix

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}

par(cex=0.6,cex.main=0.85,cex.axis=0.7)

GDSTM <- attr(deThermal,"GDSTM")
fscores <- attr(deThermal,"fscore")
fscores <- fscores[order(-fscores)]
#fscores[fscores>0]

gplots::heatmap.2(log(abs(1000*GDSTM)+1),
                  trace = "none",
#                  scale = "none",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  main = "COVID-19 Decorrelation matrix",
                  cexRow = 0.35,
                  cexCol = 0.35,
                  key.title=NA,
                  key.xlab="Log(|1000*Beta|+1)",
                  xlab="Output Feature", ylab="Input Feature")

par(op)

```

### Plots

```{r results = "asis", warning = FALSE, dpi=600, fig.height= 6.0, fig.width= 5.0}
par(mfrow=c(2,1),cex=0.8,cex.axis=0.7,cex.main=0.8)

varlist <- colnames(SalivaThermal_Avg)[!(colnames(SalivaThermal_Avg) %in% "class")]
varlist <- cbind(varlist,varlist)

univar <- uniRankVar(varlist,
	           "class~1",
	           "class",
	           SalivaThermal_Avg,
	           categorizationType = "Raw",
	           type = "LOGIT",
	           rankingTest = "AUC",
	            uniType = "Binary")

cases <- subset(SalivaThermal_Avg,class==1)
cases$class <- NULL
controls <- subset(SalivaThermal_Avg,class==0)
controls$class <- NULL
colnames(cases) <- str_remove_all(colnames(cases),"V_")
bpcases <- boxplot(cases,
                   boxwex=1.25,
                   outline = FALSE,
                   col="red",
                   las=2,
                   cex.axis=0.5,
                   ylab="Absorbance (a.u.)",
                   xlab=expression("Wavenumber (cm)"^"-1"))
#bpcontrols <- boxplot(controls,col="blue",outline = FALSE,boxwex=0.25,las=2,add = TRUE,cex.axis=0.5)
lines(apply(controls,2,mean),col="blue",lwd=3)
title("Absorbance")
legend("topright",legend=c("Cases", "Controls"),
       col=c("red", "blue"), lty=1:1,lwd=c(1,3), cex=0.6)

effectSize <- (apply(cases,2,mean)-apply(controls,2,mean))/apply(SalivaThermal_Avg[,str_detect(colnames(SalivaThermal_Avg),"V_")],2,sd)
wavenumber=as.numeric(str_remove_all(names(effectSize),"V_"))

aucvalue <- univar$orderframe$ROCAUC


wavenumber=as.numeric(str_remove_all(rownames(univar$orderframe),"V_"))
aucvalueRaw <- aucvalue[order(wavenumber)]
wavenumberRaw <- wavenumber[order(wavenumber)]
bpdiff <- plot(wavenumberRaw,aucvalueRaw,las=2,cex.axis=0.5,type="l",
                                  ylab="ROC AUC",
                    xlim = c(1300,800),
                   xlab=expression("Wavenumber (cm)"^"-1"))
title("Difference in Absorbance")



varlist <- colnames(deThermal)[!(colnames(deThermal) %in% "class")]
varlist <- cbind(varlist,varlist)

univarDe <- uniRankVar(varlist,
	           "class~1",
	           "class",
	           deThermal,
	           categorizationType = "Raw",
	           type = "LOGIT",
	           rankingTest = "AUC",
	            uniType = "Binary")


decorfeat <- colnames(deThermal)[str_detect(colnames(deThermal),"La_")]
parentfeat <- str_remove_all(decorfeat,"La_")
reldecor <- deThermal[,decorfeat]
reldecor$class <- SalivaThermal_Avg$class
cases <- subset(reldecor,class==1)
cases <- cases[,str_detect(colnames(cases),"La_")]
colnames(cases) <- str_remove_all(colnames(cases),"La_V_")
controls <- subset(reldecor,class==0)
controls <- controls[,str_detect(colnames(controls),"La_")]
colnames(controls) <- str_remove_all(colnames(controls),"La_V_")
bpcases <- boxplot(cases,
                   boxwex=2.0,
                   outline = FALSE,
                   col="red",
                   las=2,
                   cex.axis=0.5,
                   ylim=c(-0.075,0.075),
                   ylab="Absorbance (a.u.)",
                   xlab=expression("Wavenumber (cm)"^"-1"))

bpcontrols <- boxplot(controls,col="blue",ylim=c(-0.075,0.075),outline = FALSE,boxwex=1.25,las=2,add = TRUE,cex.axis=0.5)
title("Decorrelated Absorbance")
legend("topright",legend=c("Cases", "Controls"),
        fill=c("red", "blue"),
       cex=0.6)





effectSize <- (apply(cases,2,median)-apply(controls,2,median))/apply(reldecor[,str_detect(colnames(reldecor),"La_")],2,IQR)
wavenumber=as.numeric(str_remove_all(names(effectSize),"La_V_"))

aucvalue <- univarDe$orderframe$ROCAUC
depcvalue <- univarDe$orderframe$WilcoxRes.p

names(aucvalue) <- rownames(univarDe$orderframe)
plot(aucvalue[names(fscores)],fscores)


wavenumber=str_remove_all(rownames(univarDe$orderframe),"La_V_")
wavenumber=str_remove_all(wavenumber,"Ba_V_")
wavenumber=as.numeric(str_remove_all(wavenumber,"V_"))
aucvalueDe <- aucvalue[order(wavenumber)]
depcvalue <- depcvalue[order(wavenumber)]
wavenumberDe <- wavenumber[order(wavenumber)]

bpdiff <- plot(x=wavenumberDe,y=aucvalueDe,las=2,cex.axis=0.5,type="l",
                                   xlim = c(1300,800),
                                  ylab="ROCAUC",
                   xlab=expression("Wavenumber (cm)"^"-1"))
title("Difference in Decorrelated Absorbance")





par(op)

```

### Comparing Univariate ROC AUC

```{r results = "asis", warning = FALSE, dpi=600, fig.height= 5.0, fig.width= 6.0}
par(cex=0.8)
varlist <- colnames(SalivaThermal_d2)[!(colnames(SalivaThermal_d2) %in% "class")]
varlist <- cbind(varlist,varlist)
pthr <- 0.10/(ncol(SalivaThermal)-1)

univarD2 <- uniRankVar(varlist,
	           "class~1",
	           "class",
	           SalivaThermal_d2,
	           categorizationType = "Raw",
	           type = "LOGIT",
	           rankingTest = "AUC",
	            uniType = "Binary")

aucvalue <- univarD2$orderframe$ROCAUC
names(aucvalue) <- rownames(univarD2$orderframe)
pwilvalue <- univarD2$orderframe$FRes.p


wavenumber=as.numeric(str_remove_all(rownames(univarD2$orderframe),"DD_"))
aucvalueDD <- aucvalue[order(wavenumber)]
wavenumberDD <- wavenumber[order(wavenumber)]
pwilvalue <- pwilvalue[order(wavenumber)]


bpdiff <- plot(x=wavenumberRaw,y=aucvalueRaw,
               las=2,
               cex.axis=0.5,
               type="l",
               lwd=3,
                xlim=c(1300,800),
                ylim=c(0.45,0.86),
               ylab="ROC AUC",
               xlab=expression("Wavenumber (cm)"^"-1"))
title("Feature Association with COVID-19")

valroc <- pwilvalue < pthr
points(wavenumberDD[valroc],aucvalueDD[valroc],col="blue",lwd=3)

#bpdiff <- plot
valroc <- depcvalue < pthr

dc <- getDerivedCoefficients(deThermal)
for (dename in names(dc))
{
  wave <- as.numeric(str_remove(dename,"La_V_"))
  if (wave %in% wavenumberDe[valroc])
  {
    for (asfeat in names(dc[[dename]]))
    {
      points(x=as.numeric(str_remove_all(asfeat,"V_")),
               y=univarDe$orderframe[dename,"ROCAUC"],col="pink",pch=10)
    }
  }
}

points(x=wavenumberDe[valroc],
               y=aucvalueDe[valroc],col="red",pch=4,lwd=2)

legend("bottomleft",legend=c("raw", "2nd Deriv","La: Parent","La: w>0"),
       col=c("black", "blue","red","pink"), lty=c(1,0,0,0),lwd=c(3,3,2,1),pch=c(-1,1,4,10), cex=0.8)

```


### Final Table 

```{r results = "asis"}

univariate_columns <- c("caseMean","caseStd","controlMean","controlStd","controlKSP","ROCAUC","ZGLM","FRes.p")

pwilvalue <- univarDe$orderframe$FRes.p
valroc <- pwilvalue < pthr

finalTable <- univarDe$orderframe[valroc,univariate_columns]

theFormulas <- dc[rownames(finalTable)]
deFromula <- character(length(theFormulas))
names(deFromula) <- rownames(finalTable)

dx <- names(deFromula)[1]
for (dx in names(deFromula))
{
  coef <- theFormulas[[dx]]
  cname <- names(theFormulas[[dx]])
  names(cname) <- cname
  for (cf in names(coef))
  {
    if (cf != dx)
    {
      if (coef[cf]>0)
      {
        deFromula[dx] <- paste(deFromula[dx],
                               sprintf("+ %5.3f*%s",coef[cf],cname[cf]))
      }
      else
      {
        deFromula[dx] <- paste(deFromula[dx],
                               sprintf("%5.3f*%s",coef[cf],cname[cf]))
      }
    }
  }
}
orgnamez <- rownames(finalTable)
orgnamez <- str_remove_all(orgnamez,"Ba_")
orgnamez <- str_remove_all(orgnamez,"La_")
finalTable$DecorFormula <- deFromula[rownames(finalTable)]
finalTable$fscores <- fscores[rownames(finalTable)]

pander::pander(finalTable)

```


