---
title: "Univariate: Parkinson"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
  word_document: 
    reference_docx: WordStyle_FRESA.docx
    toc: yes
    fig_caption: yes
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message = FALSE,comment = "#>")
```

# PARKINSON Univariate

### Loading the libraries

```{r}
library("FRESA.CAD")
library(readxl)
op <- par(no.readonly = TRUE)
pander::panderOptions('digits', 3)
pander::panderOptions('table.split.table', 400)
pander::panderOptions('keep.trailing.zeros',TRUE)

```


## The Data

```{r}

pd_speech_features <- as.data.frame(read_excel("~/GitHub/FCA/Data/pd_speech_features.xlsx",sheet = "pd_speech_features", range = "A2:ACB758"))




```

### The Average of the Three Repetitions

Each subject had three repeated observations. Here I'll use the average of the three experiments per subject.

```{r results = "asis", warning = FALSE, dpi=600, fig.height= 6.0, fig.width= 8.0}
rep1Parkison <- subset(pd_speech_features,RID==1)
rownames(rep1Parkison) <- rep1Parkison$id
rep1Parkison$id <- NULL
rep1Parkison$RID <- NULL
rep1Parkison[,1:ncol(rep1Parkison)] <- sapply(rep1Parkison,as.numeric)

rep2Parkison <- subset(pd_speech_features,RID==2)
rownames(rep2Parkison) <- rep2Parkison$id
rep2Parkison$id <- NULL
rep2Parkison$RID <- NULL
rep2Parkison[,1:ncol(rep2Parkison)] <- sapply(rep2Parkison,as.numeric)

rep3Parkison <- subset(pd_speech_features,RID==3)
rownames(rep3Parkison) <- rep3Parkison$id
rep3Parkison$id <- NULL
rep3Parkison$RID <- NULL
rep3Parkison[,1:ncol(rep3Parkison)] <- sapply(rep3Parkison,as.numeric)

whof <- !(colnames(rep1Parkison) %in% c("gender","class"));
avgParkison <- rep1Parkison;
avgParkison[,whof] <- (rep1Parkison[,whof] + rep2Parkison[,whof] + rep3Parkison[,whof])/3
## I apply the log transform to the data
##avgParkison[,whof] <- signedlog(avgParkison[,whof])
avgParkison[,whof] <- FRESAScale(avgParkison[,whof],method="OrderLogit")$scaledData
pander::pander(table(avgParkison$class))

```

#### Standarize the names for the reporting

```{r results = "asis"}
dataframe <- avgParkison
outcome <- "class"

```



```{r results = "asis"}
pander::pander(c(rows=nrow(dataframe),col=ncol(dataframe)-1))
pander::pander(table(dataframe[,outcome]))

varlist <- colnames(dataframe)
varlist <- varlist[varlist != outcome]
varlist <- as.data.frame(cbind(name=varlist,desc=varlist))

```


## Univariate

```{r results = "asis"}

univariate_columns <- c("caseMean","caseStd","controlMean","controlStd","cohortKSP","ROCAUC","WilcoxRes.p","FRes.p")
univar <- uniRankVar(varlist,
	           paste(outcome,"~1"),
	           outcome,
	           dataframe,
	           categorizationType = "Raw",
	           type = "LOGIT",
	           rankingTest = "AUC",
	            cateGroups = c(0.1, 0.9),
	            raw.dataFrame = NULL,
	            description = ".",
	            uniType = "Binary")

pander::pander(univar$orderframe[1:20,univariate_columns])

topfiveOrg <- rownames(univar$orderframe[1:5,])

```

## Decorrelation Analysis

```{r results = "asis"}
#DEdataframe <- GDSTMDecorrelation(dataframe,Outcome=outcome,thr=0.80,verbose = TRUE,maxLoops=200,skipRelaxed=FALSE)
DEdataframe <- GDSTMDecorrelation(dataframe,thr=0.60,verbose = TRUE,maxLoops=200,skipRelaxed=FALSE)

pander::pander(c(Decorrleated_Fraction=sum(str_detect(colnames(DEdataframe),"De_"))/(ncol(DEdataframe)-1)))
pander::pander(c(Base_Fraction=sum(str_detect(colnames(DEdataframe),"Ba_"))/(ncol(DEdataframe)-1)))
demat <- attr(DEdataframe,"GDSTM")
pander::pander(c(sparse_Fraction=1.0-sum(demat==0)/ncol(demat)/nrow(demat)))
varlistDe <-  colnames(DEdataframe)[colnames(DEdataframe) != "class"];
varlistDe <- as.data.frame(cbind(name=varlistDe,desc=varlistDe))


```

### The heat maps.

```{r results = "asis", warning = FALSE, dpi=600, fig.height= 6.0, fig.width= 7.0}
cormat <- cor(dataframe[,varlist[,1]],method="spearman")
cormat[is.na(cormat)] <- 0
gplots::heatmap.2(abs(cormat),
                  trace = "none",
#                  scale = "row",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  main = "Spearman Correlation Original",
                  cexRow = 0.35,
                  cexCol = 0.35,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")


cormat <- cor(DEdataframe[,varlistDe[,1]],method="spearman")
cormat[is.na(cormat)] <- 0
gplots::heatmap.2(abs(cormat),
                  trace = "none",
#                  scale = "none",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  main = "Spearman Correlation: After GDSTM",
                  cexRow = 0.35,
                  cexCol = 0.35,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")

par(op)
```


## Univariate Decorrelated

```{r results = "asis"}

univarDe <- uniRankVar(varlistDe,
              paste(outcome,"~1"),
	            outcome,
              DEdataframe,
	            categorizationType = "Raw",
	            type = "LOGIT",
	            rankingTest = "AUC",
	            cateGroups = c(0.1, 0.9),
	            raw.dataFrame = NULL,
	            description = ".",
	            uniType = "Binary")

pander::pander(univarDe$orderframe[1:20,univariate_columns])

```

### Comparing Decorrelation vs Original

```{r results = "asis"}
topDecorNames <- rownames(univarDe$orderframe[univarDe$orderframe$FRes.p<1.0e-5,])
dc <- getDerivedCoefficients(DEdataframe)
### 2a Get only the ones that in the top features
deNames_in_dc <- topDecorNames[topDecorNames %in% names(dc)]
selectedlist <- dc[deNames_in_dc]
theDeFormulas <- selectedlist
```

### Final Table

```{r results = "asis"}
unlistdecorr <- selectedlist
names(unlistdecorr) <- NULL
unlistdecorr <- unique(names(unlist(unlistdecorr)))

finalTableDe <- univarDe$orderframe[deNames_in_dc,univariate_columns]

finalTableOr <- univar$orderframe[unique(c(topfiveOrg,unlistdecorr)),univariate_columns]

finalTable <- rbind(finalTableOr,finalTableDe)

deFromula <- character(length(theDeFormulas))
names(deFromula) <- names(theDeFormulas)

for (dx in names(deFromula))
{
  coef <- theDeFormulas[[dx]]
  cname <- names(theDeFormulas[[dx]])
  names(cname) <- cname
  for (cf in names(coef))
  {
    if (cf != dx)
    {
      if (coef[cf]>0)
      {
        deFromula[dx] <- paste(deFromula[dx],
                               sprintf("+ %5.3f*%s",coef[cf],cname[cf]))
      }
      else
      {
        deFromula[dx] <- paste(deFromula[dx],
                               sprintf("%5.3f*%s",coef[cf],cname[cf]))
      }
    }
  }
}
orgnamez <- rownames(finalTable)
orgnamez <- str_remove_all(orgnamez,"Ba_")
orgnamez <- str_remove_all(orgnamez,"De_")
finalTable$uAUC <- univar$orderframe[orgnamez,"ROCAUC"]
finalTable$DecorFormula <- deFromula[rownames(finalTable)]
finalTable <- finalTable[order(-finalTable$ROCAUC),]
pander::pander(finalTable)
```

