---
title: "TADPOLE_Survival"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
  word_document: 
    reference_docx: WordStyle_FRESA.docx
    toc: yes
    fig_caption: yes
editor_options:
  chunk_output_type: console
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message = FALSE,comment = "#>")
```

# TADPOLE Univariate

### The run options
```{r}
TrainFraction <- 0.50;
sig_pvalue <- 0.01
thr <- c(0.05,0.2,0.4,0.6,0.8,0.95);
method  <- c("fast","pearson","spearman","kendall");
type <- c("LM","RLM")
DeOutcome <- c("NoDriven","Driven")
corRankS <- c("FALSE","TRUE")

thr <- thr[1]
method  <- method[1]
type <- type[1] 
DeOutcome <- DeOutcome[1]
corRankS <- corRankS[2]


corRank <- corRankS == "TRUE"

```


### Loading the libraries

```{r}
library("FRESA.CAD")
library(readxl)
library("whitening")
library("fpc")
library("robust")
op <- par(no.readonly = TRUE)
pander::panderOptions('digits', 3)
pander::panderOptions('table.split.table', 400)
pander::panderOptions('keep.trailing.zeros',TRUE)

```
## The data set

```{r}
TADPOLE_D1_D2 <- read.csv("~/GitHub/BSWiMS/Data/TADPOLE/TADPOLE_D1_D2.csv")
TADPOLE_D1_D2_Dict <- read.csv("~/GitHub/BSWiMS/Data/TADPOLE/TADPOLE_D1_D2_Dict.csv")
TADPOLE_D1_D2_Dict_LR <- as.data.frame(read_excel("~/GitHub/BSWiMS/Data/TADPOLE/TADPOLE_D1_D2_Dict_LR.xlsx",sheet = "LeftRightFeatures"))


rownames(TADPOLE_D1_D2_Dict) <- TADPOLE_D1_D2_Dict$FLDNAME

```

## Conditioning the data

```{r results = "asis", warning = FALSE, dpi=600, fig.height= 6.0, fig.width= 8.0}

# mm3 to mm
isVolume <- c("Ventricles","Hippocampus","WholeBrain","Entorhinal","Fusiform","MidTemp","ICV",
              TADPOLE_D1_D2_Dict$FLDNAME[str_detect(TADPOLE_D1_D2_Dict$TEXT,"Volume")]
              )


#TADPOLE_D1_D2[,isVolume] <- apply(TADPOLE_D1_D2[,isVolume],2,'^',(1/3))
TADPOLE_D1_D2[,isVolume] <- TADPOLE_D1_D2[,isVolume]^(1/3)

# mm2 to mm
isArea <- TADPOLE_D1_D2_Dict$FLDNAME[str_detect(TADPOLE_D1_D2_Dict$TEXT,"Area")]
TADPOLE_D1_D2[,isArea] <- sqrt(TADPOLE_D1_D2[,isArea])

# Get only cross sectional measurements
FreeSurfersetCross <- str_detect(colnames(TADPOLE_D1_D2),"UCSFFSX")

# The subset of baseline measurements
baselineTadpole <- subset(TADPOLE_D1_D2,VISCODE=="bl")
table(baselineTadpole$DX)
table(baselineTadpole$DX_bl)

rownames(baselineTadpole) <- baselineTadpole$PTID


validBaselineTadpole <- cbind(DX=baselineTadpole$DX_bl,
                                 AGE=baselineTadpole$AGE,
                                 Gender=1*(baselineTadpole$PTGENDER=="Female"),
                                 ADAS11=baselineTadpole$ADAS11,
                                 ADAS13=baselineTadpole$ADAS13,
                                 MMSE=baselineTadpole$MMSE,
                                 RAVLT_immediate=baselineTadpole$RAVLT_immediate,
                                 RAVLT_learning=baselineTadpole$RAVLT_learning,
                                 RAVLT_forgetting=baselineTadpole$RAVLT_forgetting,
                                 RAVLT_perc_forgetting=baselineTadpole$RAVLT_perc_forgetting,
                                 FAQ=baselineTadpole$FAQ,
                                 Ventricles=baselineTadpole$Ventricles,
                                 Hippocampus=baselineTadpole$Hippocampus,
                                 WholeBrain=baselineTadpole$WholeBrain,
                                 Entorhinal=baselineTadpole$Entorhinal,
                                 Fusiform=baselineTadpole$Fusiform,
                                 MidTemp=baselineTadpole$MidTemp,
                                 ICV=baselineTadpole$ICV,
                                 baselineTadpole[,FreeSurfersetCross])


LeftFields <- TADPOLE_D1_D2_Dict_LR$LFN
names(LeftFields) <- LeftFields
LeftFields <- LeftFields[LeftFields %in% colnames(validBaselineTadpole)]
RightFields <- TADPOLE_D1_D2_Dict_LR$RFN
names(RightFields) <- RightFields
RightFields <- RightFields[RightFields %in% colnames(validBaselineTadpole)]

## Normalize to ICV
validBaselineTadpole$Ventricles=validBaselineTadpole$Ventricles/validBaselineTadpole$ICV
validBaselineTadpole$Hippocampus=validBaselineTadpole$Hippocampus/validBaselineTadpole$ICV
validBaselineTadpole$WholeBrain=validBaselineTadpole$WholeBrain/validBaselineTadpole$ICV
validBaselineTadpole$Entorhinal=validBaselineTadpole$Entorhinal/validBaselineTadpole$ICV
validBaselineTadpole$Fusiform=validBaselineTadpole$Fusiform/validBaselineTadpole$ICV
validBaselineTadpole$MidTemp=validBaselineTadpole$MidTemp/validBaselineTadpole$ICV

leftData <- validBaselineTadpole[,LeftFields]/validBaselineTadpole$ICV
RightData <- validBaselineTadpole[,RightFields]/validBaselineTadpole$ICV

## get mean and relative difference 
meanLeftRight <- (leftData + RightData)/2
difLeftRight <- abs(leftData - RightData)
reldifLeftRight <- difLeftRight/meanLeftRight
colnames(meanLeftRight) <- paste("M",colnames(meanLeftRight),sep="_")
colnames(difLeftRight) <- paste("D",colnames(difLeftRight),sep="_")
colnames(reldifLeftRight) <- paste("RD",colnames(reldifLeftRight),sep="_")


validBaselineTadpole <- validBaselineTadpole[,!(colnames(validBaselineTadpole) %in% 
                                               c(LeftFields,RightFields))]
#validBaselineTadpole <- cbind(validBaselineTadpole,meanLeftRight,difLeftRight,reldifLeftRight)
validBaselineTadpole <- cbind(validBaselineTadpole,meanLeftRight,difLeftRight)

## Remove columns with too many NA more than %15 of NA
nacount <- apply(is.na(validBaselineTadpole),2,sum)/nrow(validBaselineTadpole) < 0.15
diagnose <- validBaselineTadpole$DX
pander::pander(table(diagnose))
validBaselineTadpole <- validBaselineTadpole[,nacount]
## Remove character columns
ischar <- sapply(validBaselineTadpole,class) == "character"
validBaselineTadpole <- validBaselineTadpole[,!ischar]
## Place back diagnose
validBaselineTadpole$DX <- diagnose


validBaselineTadpole <- validBaselineTadpole[complete.cases(validBaselineTadpole),]
ischar <- sapply(validBaselineTadpole,class) == "character"
validBaselineTadpole[,!ischar] <- sapply(validBaselineTadpole[,!ischar],as.numeric)

colnames(validBaselineTadpole) <- str_remove_all(colnames(validBaselineTadpole),"_UCSFFSX_11_02_15_UCSFFSX51_08_01_16")
colnames(validBaselineTadpole) <- str_replace_all(colnames(validBaselineTadpole)," ","_")
validBaselineTadpole$LONISID <- NULL
validBaselineTadpole$IMAGEUID <- NULL
validBaselineTadpole$LONIUID <- NULL

diagnose <- as.character(validBaselineTadpole$DX)
validBaselineTadpole$DX <- diagnose
pander::pander(table(validBaselineTadpole$DX))


validBaselineTadpole[validBaselineTadpole$DX %in% c("EMCI","LMCI"),"DX"] <- "MCI" 
validBaselineTadpole[validBaselineTadpole$DX %in% c("CN","SMC"),"DX"] <- "NL" 

pander::pander(table(validBaselineTadpole$DX))



```

## Get the Time To Event on MCI Subjects

```{r results = "asis", warning = FALSE, dpi=600, fig.height= 6.0, fig.width= 8.0}

subjectsID <- rownames(validBaselineTadpole)
visitsID <- unique(TADPOLE_D1_D2$VISCODE)
baseDx <- TADPOLE_D1_D2[TADPOLE_D1_D2$VISCODE=="bl",c("PTID","DX","EXAMDATE")]
rownames(baseDx) <- baseDx$PTID 
baseDx <- baseDx[subjectsID,]
lastDx <- baseDx
toDementia <- baseDx
table(lastDx$DX)
hasDementia <- lastDx$PTID[str_detect(lastDx$DX,"Dementia")]


for (vid in visitsID)
{
  DxValue <- TADPOLE_D1_D2[TADPOLE_D1_D2$VISCODE==vid,c("PTID","DX","EXAMDATE")]
  rownames(DxValue) <- DxValue$PTID 
  DxValue <- DxValue[DxValue$PTID %in% subjectsID,]
  noDX <- DxValue$PTID[nchar(DxValue$DX) < 1]
  print(length(noDX))
  DxValue[noDX,] <- lastDx[noDX,]
  inLast <- lastDx$PTID[lastDx$PTID %in% DxValue$PTID]
  print(length(inLast))
  lastDx[inLast,] <- DxValue[inLast,]
  noDementia <- !(toDementia$PTID %in% hasDementia)
  toDementia[noDementia,] <- lastDx[noDementia,]
  hasDementia <- unique(c(hasDementia,lastDx$PTID[str_detect(lastDx$DX,"Dementia")]))
}
table(lastDx$DX)
baseMCI <-baseDx$PTID[baseDx$DX == "MCI"]
lastDementia <- lastDx$PTID[str_detect(lastDx$DX,"Dementia")]
lastDementia2 <- toDementia$PTID[str_detect(toDementia$DX,"Dementia")]
lastNL <- lastDx$PTID[str_detect(lastDx$DX,"NL")]

MCIatBaseline <- baseDx[baseMCI,]
MCIatEvent <- toDementia[baseMCI,]
MCIatLast <- lastDx[baseMCI,]

MCIconverters <- MCIatBaseline[baseMCI %in% lastDementia,]
MCI_No_converters <- MCIatBaseline[!(baseMCI %in% MCIconverters$PTID),]
MCIconverters$TimeToEvent <- (as.Date(toDementia[MCIconverters$PTID,"EXAMDATE"]) 
                                   - as.Date(MCIconverters$EXAMDATE))

sum(MCIconverters$TimeToEvent ==0)


MCIconverters$AtEventDX <- MCIatEvent[MCIconverters$PTID,"DX"]
MCIconverters$LastDX <- MCIatLast[MCIconverters$PTID,"DX"]

MCI_No_converters$TimeToEvent <- (as.Date(lastDx[MCI_No_converters$PTID,"EXAMDATE"]) 
                                   - as.Date(MCI_No_converters$EXAMDATE))

MCI_No_converters$LastDX <- MCIatLast[MCI_No_converters$PTID,"DX"]

MCI_No_converters <- subset(MCI_No_converters,TimeToEvent > 0)



```


# Prognosis MCI to AD Conversion

## the set
```{r results = "asis", warning = FALSE, dpi=600, fig.height= 6.0, fig.width= 8.0}

MCIPrognosisIDs <- c(MCIconverters$PTID,MCI_No_converters$PTID)

TADPOLECrossMRI <- validBaselineTadpole[MCIPrognosisIDs,]
table(TADPOLECrossMRI$DX)
TADPOLECrossMRI$DX <- NULL
TADPOLECrossMRI$status <- 1*(rownames(TADPOLECrossMRI) %in% MCIconverters$PTID)
table(TADPOLECrossMRI$status)
TADPOLECrossMRI$TimeToEvent <- numeric(nrow(TADPOLECrossMRI))
TADPOLECrossMRI[MCIconverters$PTID,"TimeToEvent"] <- MCIconverters$TimeToEvent
TADPOLECrossMRI[MCI_No_converters$PTID,"TimeToEvent"] <- MCI_No_converters$TimeToEvent



set.seed(1)
trainCases <- sample(nrow(TADPOLECrossMRI),nrow(TADPOLECrossMRI)*TrainFraction)

TADPOLE_Conv_TRAIN <- TADPOLECrossMRI[trainCases,]
TADPOLE_Conv_TEST <- TADPOLECrossMRI[-trainCases,]

pander::pander(table(TADPOLE_Conv_TRAIN$status))
pander::pander(table(TADPOLE_Conv_TEST$status))

surTimeTrain <- TADPOLE_Conv_TRAIN$TimeToEvent
surTimeTest <- TADPOLE_Conv_TEST$TimeToEvent


TADPOLE_Conv_TRAIN$TimeToEvent <- NULL
TADPOLE_Conv_TEST$TimeToEvent <- NULL

```



#### Analysis parameters

```{r results = "asis"}
dataframe <- TADPOLE_Conv_TRAIN
dataframeTest <- TADPOLE_Conv_TEST
outcome <- "status"


univariate_columns <- c("caseMean","caseStd","controlMean","controlStd","controlKSP","cStatCorr","ZGLM","pvalue")


```

## The features to analyze

```{r results = "asis"}
pander::pander(c(rows=nrow(dataframe),col=ncol(dataframe)-1))
pander::pander(table(dataframe[,outcome]))

varlist <- colnames(dataframe)
varlist <- varlist[!(varlist %in% c(outcome,"TimeToEvent"))]
varlist <- as.data.frame(cbind(name=varlist,desc=varlist))

```


## Univariate RAW

```{r results = "asis"}

dataframeTime <- cbind(dataframe,SurvTime=surTimeTrain)


univarTrain <- uniRankVar(varlist,
	           paste("Surv(SurvTime,",outcome,")~1"),
	           outcome,
	           dataframeTime,
	           categorizationType = "Raw",
	           type = "COX",
	           rankingTest = "Ztest",
	            uniType = "Binary",
	           timeOutcome="SurvTime")

univarTrain$orderframe$pvalue=pnorm(abs(univarTrain$orderframe$ZGLM),lower.tail = FALSE)

pander::pander(univarTrain$orderframe[1:10,univariate_columns])

dataframeTime <- cbind(dataframeTest,SurvTime=surTimeTest)


univar <- uniRankVar(varlist,
	           paste("Surv(SurvTime,",outcome,")~1"),
	           outcome,
	           dataframeTime,
	           categorizationType = "Raw",
	           type = "COX",
	           rankingTest = "Ztest",
	            uniType = "Binary",
	           timeOutcome="SurvTime")

univar$orderframe$pvalue=pnorm(abs(univar$orderframe$ZGLM),lower.tail = FALSE)

```


===================================================================================================================

## Decorrelation Analysis 

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 4.0, fig.width= 6.0}

if (DeOutcome != "Driven")
{
  DEdataframeTrain <- GDSTMDecorrelation(dataframe,thr=thr,method=method,type=type,corRank=corRank,relaxed=FALSE,verbose = TRUE)
} else
{
  DEdataframeTrain <- GDSTMDecorrelation(dataframe,Outcome=outcome,thr=thr,method=method,type=type,relaxed=FALSE,corRank=corRank,verbose = TRUE)
}

#cormat <- abs(cor(DEdataframeTrain[,colnames(DEdataframeTrain) != outcome],method="spearman"));
cormat <- abs(cor(DEdataframeTrain[,colnames(DEdataframeTrain) != outcome],method="pearson"));
diag(cormat) <- 0;

gplots::heatmap.2(abs(cormat),
                  trace = "none",
#                  scale = "row",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  main = "Correlation",
                  cexRow = 0.35,
                  cexCol = 0.35,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")

pander::pander(max(cormat))
maxcor <- apply(cormat,2,max)
#pander::pander(maxcor[maxcor>thr])

demat <- attr(DEdataframeTrain,"GDSTM")
DEdataframe <- predictDecorrelate(DEdataframeTrain,dataframeTest)


pander::pander(c(Cols=ncol(demat),Rows=nrow(demat)))
totFe <- ncol(DEdataframe)-1
totBa <- sum(str_detect(colnames(DEdataframe),"Ba_"))
totDe <- sum(str_detect(colnames(DEdataframe),"De_"))
toUnmat <- sum(apply(demat!=0,2,sum)==1)
unaltered <- totFe - ncol(demat) + toUnmat
pander::pander(c(Features=totFe,totBa=totBa,totDe=totDe,unaltered=totFe-totBa-totDe,totuna=unaltered))
pander::pander(c(Decorrleated_Fraction=sum(str_detect(colnames(DEdataframe),"De_"))/(ncol(DEdataframe)-1)))
pander::pander(c(Base_Fraction=sum(str_detect(colnames(DEdataframe),"Ba_"))/(ncol(DEdataframe)-1)))

pander::pander(c(Unaltered_Fraction=unaltered/totFe))
pander::pander(c(sparcity=(totFe-ncol(demat)+sum(abs(demat)!=0))/totFe/totFe))

varlistDe <-  colnames(demat)[apply(demat!=0,2,sum)>1];
varlistDe <- as.data.frame(cbind(name=varlistDe,desc=varlistDe))


```


## Univariate Decorrelated 

```{r results = "asis"}
varlist_DeAll <- colnames(DEdataframe)
varlist_DeAll <- varlist_DeAll[!(varlist_DeAll %in% c(outcome,"TimeToEvent"))]
varlist_DeAll <- as.data.frame(cbind(name=varlist_DeAll,desc=varlist_DeAll))

DEdataframeTimeTrain <- cbind(DEdataframeTrain,SurvTime=surTimeTrain)

#DEdataframe$surTime <- surTime

univarDeAllTrain <- uniRankVar(varlist_DeAll,
	           paste("Surv(SurvTime,",outcome,")~1"),
	           outcome,
	           DEdataframeTimeTrain,
	           categorizationType = "Raw",
	           type = "COX",
	           rankingTest = "Ztest",
	            uniType = "Binary",
	           timeOutcome="SurvTime")

univarDeAllTrain$orderframe$pvalue=pnorm(abs(univarDeAllTrain$orderframe$ZGLM),lower.tail = FALSE)

pander::pander(univarDeAllTrain$orderframe[1:10,univariate_columns])

univarDeTrain <- univarDeAllTrain
univarDeTrain$orderframe <- univarDeAllTrain$orderframe[varlistDe[,1],]
univarDeTrain$orderframe <- univarDeTrain$orderframe[order(univarDeTrain$orderframe$pvalue),]

pander::pander(univarDeTrain$orderframe[1:10,univariate_columns])



```


### Survival plot of top Train feature 

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 6.0, fig.width= 9.0}
par(cex=0.7)
s = Surv(DEdataframeTimeTrain$SurvTime, DEdataframeTimeTrain$status)
topvar <- univarDeTrain$orderframe$Name[1]
cmod <- coxph(formula(paste("Surv(SurvTime, status)~",topvar)),DEdataframeTimeTrain);
pmod <- predict(cmod,DEdataframeTimeTrain,type="lp")
pmor <- predict(cmod,DEdataframeTimeTrain,type="risk")
plot(survfit(s ~ pmod>0),main=topvar, xlab = "Years", ylab = "Probability of survival", lty=c(2,1), lwd=2)
legend("topright", legend = c("High Risk", "Low Risk"), lty =c(1,2), lwd=2)
pLR=EmpiricalSurvDiff(DEdataframeTimeTrain$SurvTime,DEdataframeTimeTrain$status,pmod>0)
pander::pander(pLR$pvalue)

perdsurvTest <- cbind(DEdataframeTimeTrain$SurvTime,
                  DEdataframeTimeTrain$status,
                  pmod,
                  pmor)
prSurvTest <- predictionStats_survival(perdsurvTest,topvar,cex=0.75)

pander::pander(survdiff(s ~ pmod>0))

```

### Test Univariate 
```{r}
DEdataframeTime <- cbind(DEdataframe,SurvTime=surTimeTest)

#DEdataframe$surTime <- surTime



univarDetestAll <- uniRankVar(varlist_DeAll,
	           paste("Surv(SurvTime,",outcome,")~1"),
	           outcome,
	           DEdataframeTime,
	           categorizationType = "Raw",
	           type = "COX",
	           rankingTest = "Ztest",
	            uniType = "Binary",
	           timeOutcome="SurvTime")


univarDetestAll$orderframe$pvalue=pnorm(abs(univarDetestAll$orderframe$ZGLM),lower.tail = FALSE)

pander::pander(univarDetestAll$orderframe[1:20,univariate_columns])

univarDetest <- univarDetestAll
univarDetest$orderframe <- univarDetest$orderframe[varlistDe[,1],]
univarDetest$orderframe <- univarDetest$orderframe[order(univarDetest$orderframe$pvalue),]
pander::pander(univarDetest$orderframe[1:10,univariate_columns])


```

### Survival plot of top feature 

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 6.0, fig.width= 9.0}
par(cex=0.7)
s = Surv(DEdataframeTime$SurvTime, DEdataframeTime$status)
topvar <- univarDeTrain$orderframe$Name[1]
cmod <- coxph(formula(paste("Surv(SurvTime, status)~",topvar)),DEdataframeTimeTrain);
pmod <- predict(cmod,DEdataframeTime,type="lp")
pmor <- predict(cmod,DEdataframeTime,type="risk")
plot(survfit(s ~ pmod>0),main=topvar, xlab = "Years", ylab = "Probability of survival", lty=c(2,1), lwd=2)
legend("topright", legend = c("High Risk", "Low Risk"), lty =c(1,2), lwd=2)
pLR=EmpiricalSurvDiff(DEdataframeTime$SurvTime,DEdataframeTime$status,pmod>0)
pander::pander(pLR$pvalue)

perdsurvTest <- cbind(DEdataframeTime$SurvTime,
                  DEdataframeTime$status,
                  pmod,
                  pmor)
prSurvTest <- predictionStats_survival(perdsurvTest,topvar,cex=0.75)

pander::pander(survdiff(s ~ pmod>0))

```

### The Top Significant Figures 

```{r results = "asis"}
pthr <- sig_pvalue/(ncol(dataframe)-1)

topDecorNames <- rownames(univarDeTrain$orderframe[univarDeTrain$orderframe$pvalue<pthr,])
topDecorNames <- unique(c(topDecorNames,rownames(univarDeTrain$orderframe[1:5,])))

dc <- getDerivedCoefficients(DEdataframeTrain)
### 2a Get only the ones that in the top features
deNames_in_dc <- topDecorNames[topDecorNames %in% names(dc)]
selectedlist <- dc[deNames_in_dc]
theDeFormulas <- selectedlist


pander::pander(c(Average_Latent_Size=length(unlist(theDeFormulas))/length(theDeFormulas)))
```


### Final Table 

```{r results = "asis"}
unlistdecorr <- selectedlist
names(unlistdecorr) <- NULL
unlistdecorr <- unique(names(unlist(unlistdecorr)))

finalTableDe <- univarDeTrain$orderframe[deNames_in_dc,univariate_columns]
finalTableDetest <- univarDetest$orderframe[deNames_in_dc,univariate_columns]

finalTable <- finalTableDetest

deFromula <- character(length(theDeFormulas))
names(deFromula) <- names(theDeFormulas)

for (dx in names(deFromula))
{
  coef <- theDeFormulas[[dx]]
  cname <- names(theDeFormulas[[dx]])
  names(cname) <- cname
  for (cf in names(coef))
  {
    if (cf != dx)
    {
      if (coef[cf]>0)
      {
        deFromula[dx] <- paste(deFromula[dx],
                               sprintf("+ %5.3f*%s",coef[cf],cname[cf]))
      }
      else
      {
        deFromula[dx] <- paste(deFromula[dx],
                               sprintf("%5.3f*%s",coef[cf],cname[cf]))
      }
    }
  }
}
orgnamez <- rownames(finalTable)
orgnamez <- str_remove_all(orgnamez,"Ba_")
orgnamez <- str_remove_all(orgnamez,"De_")
finalTable$Train_ZGLM <- finalTableDe[rownames(finalTable),"ZGLM"]
finalTable$Train_cStatCorr <- finalTableDe[rownames(finalTable),"cStatCorr"]
finalTable$pvalue <- pnorm(abs(finalTable$ZGLM),lower.tail = FALSE)
finalTable$uZGLM <- univar$orderframe[orgnamez,"ZGLM"]
finalTable$DecorFormula <- deFromula[rownames(finalTable)]
finalTable <- finalTable[order(finalTable$pvalue),]
finalTable$ROCAUC <- NULL
finalTable$kendall.p <- NULL
pander::pander(finalTable)

save(finalTable,file=paste("FinalTable",thr,method,type,DeOutcome,corRankS,"TRUE.RDATA",sep="_"))

```



### The bhattacharyya between the two classes: TRAIN



```{r results = "asis", warning = FALSE, dpi=600, fig.height= 4.0, fig.width= 8.0}

par(op)
par(mfrow=c(1,2),cex=0.9)

pvals <- univarTrain$orderframe$pvalue
names(pvals) <- rownames(univarTrain$orderframe)
praw <- p.adjust(pvals,method="BH")
topSigRaw <- names(praw[praw<sig_pvalue])


pvalsDe <- univarDeAllTrain$orderframe$pvalue
names(pvalsDe) <- rownames(univarDeAllTrain$orderframe)
pDe <- p.adjust(pvalsDe,method="BH")
topSigDe <- names(pDe[pDe<sig_pvalue])

pDeDe <- pDe[names(pDe) %in% varlistDe[,1]]
pDeDe <- pDeDe[pDeDe<sig_pvalue]

pander::pander(c(Raw_Significant=length(topSigRaw),De_Significant=length(topSigDe),SigDe=length(pDeDe)))
sig_p <- min(c(sig_pvalue,1.0e4*praw[10]))

praw <- praw[praw <= sig_p]

pander::pander(c(Raw_Significant=length(praw),Pvalue=sig_p))
pander::pander(as.matrix(praw))
if (sum(pDe <= sig_p)>1)
{
  pDe <- pDe[pDe <= sig_p]
} else
{
  pDe <- pDe[1:2]
}

topvarraw <- names(praw)


subcases <- subset(dataframe,get(outcome)==1)
subcontrols <- subset(dataframe,get(outcome)==0)


casesCov <- MASS::cov.rob(subcases[,topvarraw])
meanCases <- casesCov$center
covCases <- casesCov$cov

controlsCov <- MASS::cov.rob(subcontrols[,topvarraw])
meanControls <- controlsCov$center
covControls <- controlsCov$cov


distRaw <- bhattacharyya.dist(meanCases, meanControls, covCases, covControls)
maha1 <- mahalanobis(meanCases,meanControls,covControls)
maha2 <- mahalanobis(meanControls,meanCases,covCases)
mahaDisRaw <- sqrt(0.5*(maha1+maha2))

disttoControl <- mahalanobis(dataframeTest[,topvarraw],meanControls,covControls)
disttoCases <- mahalanobis(dataframeTest[,topvarraw],meanCases,covCases)
pbMaha <- predictionStats_binary(cbind(dataframeTest[,outcome],disttoControl-disttoCases),"Raw Maha:ROC",cex=0.8)
#mlNB <- NAIVE_BAYES(formula(paste(outcome,"~.")),dataframe[,c(outcome,topvarraw)],pca=FALSE)
#pb <- predictionStats_binary(cbind(dataframeTest[,outcome],predict(mlNB,dataframeTest[,c(outcome,topvarraw)])),"Raw NB:ROC")
#mlSig <- CVsignature(formula(paste(outcome,"~.")),dataframe[,c(outcome,topvarraw)])
#pb <- predictionStats_binary(cbind(dataframeTest[,outcome],predict(mlSig,dataframeTest[,c(outcome,topvarraw)])),"Raw Sig:ROC")

### Decorrelated Analysis

subcases <- subset(DEdataframeTrain,get(outcome)==1)
subcontrols <- subset(DEdataframeTrain,get(outcome)==0)


topvarDe <- names(pDe)
pander::pander(c(Decorrelated_Significant=length(pDe)))

pander::pander(as.matrix(pDe))
casesCov <- MASS::cov.rob(subcases[,topvarDe])
meanCases <- casesCov$center
covCases <- casesCov$cov

controlsCov <- MASS::cov.rob(subcontrols[,topvarDe])
meanControls <- controlsCov$center
covControls <- controlsCov$cov


distDe <- bhattacharyya.dist(meanCases, meanControls, covCases, covControls)

maha1 <- mahalanobis(meanCases,meanControls,covControls)
maha2 <- mahalanobis(meanControls,meanCases,covCases)
mahaDisDe <- sqrt(0.5*(maha1+maha2))


disttoControl <- mahalanobis(DEdataframe[,topvarDe],meanControls,covControls)
disttoCases <- mahalanobis(DEdataframe[,topvarDe],meanCases,covCases)
pbMaha <- predictionStats_binary(cbind(DEdataframe[,outcome],disttoControl-disttoCases),"De Maha:ROC",cex=0.8)



mlKNNRaw <- KNN_method(formula(paste(outcome,"~.")),dataframe[,c(outcome,topSigRaw)])
pbKNN <- predictionStats_binary(cbind(dataframeTest[,outcome],predict(mlKNNRaw,dataframeTest[,c(outcome,topSigRaw)])),"Raw KNN:ROC",cex=0.8)

pander::pander(pbKNN$aucs)
pander::pander(pbMaha$aucs)

mlKNN <- KNN_method(formula(paste(outcome,"~.")),DEdataframeTrain[,c(outcome,topSigDe)])
pbKNN <- predictionStats_binary(cbind(dataframeTest[,outcome],predict(mlKNN,DEdataframe[,c(outcome,topSigDe)])),"De KNN:ROC",cex=0.8)

pander::pander(c(Raw_Bhattacharyya=distRaw,Raw_Mahalanobis=mahaDisRaw))
pander::pander(c(De_Bhattacharyya=distDe,De_Mahalanobis=mahaDisDe))
pander::pander(pbKNN$aucs)
pander::pander(pbMaha$aucs)

```

### Comparing the ROC KNN Train Test curves

```{r results = "asis", warning = FALSE, dpi=600, fig.height= 4.0, fig.width= 8.0}
par(op)
par(mfrow=c(1,2),cex=0.8)

pbKNNTrain <- predictionStats_binary(cbind(dataframe[,outcome],predict(mlKNN,DEdataframeTrain[,c(outcome,topSigDe)])),"De Train KNN:ROC",cex=0.8)
pbKNNTest <- predictionStats_binary(cbind(dataframeTest[,outcome],predict(mlKNN,DEdataframe[,c(outcome,topSigDe)])),"De Test KNN:ROC",cex=0.8)

```

