---
title: "IDeA, EFA: Cross Validation Parkinson's"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
  word_document: 
    reference_docx: WordStyle_FRESA.docx
    toc: yes
    fig_caption: yes
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message = FALSE,comment = "#>")
```

# Parkinson's EFA-IDeA

### Loading the libraries

```{r}
library("FRESA.CAD")
library(psych)
library(whitening)
library("vioplot")

library(readxl)
op <- par(no.readonly = TRUE)
pander::panderOptions('digits', 3)
pander::panderOptions('table.split.table', 400)
pander::panderOptions('keep.trailing.zeros',TRUE)

```


## The Data

```{r}

pd_speech_features <- as.data.frame(read_excel("~/GitHub/FCA/Data/pd_speech_features.xlsx",sheet = "pd_speech_features", range = "A2:ACB758"))




```

### The Average of the Three Repetitions

Each subject had three repeated observations. Here I'll use the average of the three experiments per subject.

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 6.0, fig.width= 8.0}
rep1Parkison <- subset(pd_speech_features,RID==1)
rownames(rep1Parkison) <- rep1Parkison$id
rep1Parkison$id <- NULL
rep1Parkison$RID <- NULL
rep1Parkison[,1:ncol(rep1Parkison)] <- sapply(rep1Parkison,as.numeric)

rep2Parkison <- subset(pd_speech_features,RID==2)
rownames(rep2Parkison) <- rep2Parkison$id
rep2Parkison$id <- NULL
rep2Parkison$RID <- NULL
rep2Parkison[,1:ncol(rep2Parkison)] <- sapply(rep2Parkison,as.numeric)

rep3Parkison <- subset(pd_speech_features,RID==3)
rownames(rep3Parkison) <- rep3Parkison$id
rep3Parkison$id <- NULL
rep3Parkison$RID <- NULL
rep3Parkison[,1:ncol(rep3Parkison)] <- sapply(rep3Parkison,as.numeric)

whof <- !(colnames(rep1Parkison) %in% c("gender","class"));
avgParkison <- rep1Parkison;
avgParkison[,whof] <- (rep1Parkison[,whof] + rep2Parkison[,whof] + rep3Parkison[,whof])/3

pander::pander(table(avgParkison$class))




```

#### Training and testing sets

```{r results = "asis"}
dataframe <- avgParkison
outcome <- "class"

trainFraction <- 0.5
cvloops <- 30
rhoThreshold <- 0.6
nEFAFactors <- 20
topFeat <- 20

aucTHR <- 0.60

set.seed(10)
colors=c("red","blue","darkblue","green","darkgreen")

```


### Data specs

```{r results = "asis"}
pander::pander(c(rows=nrow(dataframe),col=ncol(dataframe)-1))
pander::pander(table(dataframe[,outcome]))


```




## Shrinkage Analysis

```{r results = "asis"}

AUCthrlist <- numeric()
RawAUCs <- numeric()
IDeAAUCs <- numeric()
IDeAAUCs_LV <- numeric()
EFAAUCs <- numeric()
EFAAUCs_LV <- numeric()

RawBACs <- numeric()
IDeABACs <- numeric()
IDeABACs_LV <- numeric()
EFABACs <- numeric()
EFABACs_LV <- numeric()

ChangeRawAUCs <- numeric()
ChangeIDeAAUCs <- numeric()
ChangeIDeAAUCs_LV <- numeric()
ChangeEFAAUCs <- numeric()
ChangeEFAAUCs_LV <- numeric()

ChangeRawBACs <- numeric()
ChangeIDeABACs <- numeric()
ChangeIDeABACs_LV <- numeric()
ChangeEFABACs <- numeric()
ChangeEFABACs_LV <- numeric()


allposRaw <- numeric()
allposIDeA <- numeric()
allposIDeA_LV <- numeric()
allposEFA <- numeric()
allposEFA_LV <- numeric()

cvloop <- 1
varlist <- colnames(dataframe)
varlist <- varlist[varlist != outcome]
for (cvloop in c(1:cvloops))
{
    trainSample <- sample(nrow(dataframe),nrow(dataframe)*trainFraction)
    
    trainDataFrame <- dataframe[trainSample,]
    testDataFrame <- dataframe[-trainSample,]
    
    univar <- uniRankVar(varlist,
    	           paste(outcome,"~1"),
    	           outcome,
    	           trainDataFrame,
                  rankingTest = "CStat")
    
    
    univarTest <- uniRankVar(varlist,
    	           paste(outcome,"~1"),
    	           outcome,
    	           trainDataFrame,
    	           testData=testDataFrame,
                  rankingTest = "CStat")
    
    topfeat <- c(1:nrow(univar$orderframe)) <= topFeat
    posdisc <- (univar$orderframe$ROCAUC >= aucTHR) & topfeat
    RAWaucTHR <- min(univar$orderframe[posdisc,"ROCAUC"])
    AUCthrlist <- c(AUCthrlist,RAWaucTHR)
    postestrain <- posdisc & (univarTest$orderframe$ROCAUC >= aucTHR)
    allposRaw <- c(allposRaw,sum(postestrain)/(0.00001+sum(posdisc)))
    
    trainBACC <- (univar$orderframe$Sensitivity[posdisc] + univar$orderframe$Specificity[posdisc])/2.0
    testBACC <- (univarTest$orderframe$Sensitivity[posdisc] + univarTest$orderframe$Specificity[posdisc])/2.0
    ChangeRawAUCs <- c(ChangeRawAUCs,univarTest$orderframe$ROCAUC[posdisc] - univar$orderframe$ROCAUC[posdisc])
    ChangeRawBACs <- c(ChangeRawBACs,testBACC - trainBACC)
    
    RawAUCs <- c(RawAUCs,univarTest$orderframe$ROCAUC[posdisc])
    RawBACs <- c(RawBACs,testBACC[posdisc])
    
    #################################### IDeA Pearson
    
    DEdataframe <- IDeA(trainDataFrame,thr=rhoThreshold)
    predTestDe <- predictDecorrelate(DEdataframe,testDataFrame)
    
    
    varlistDe <-  colnames(DEdataframe)[colnames(DEdataframe) != outcome];
    univarDe <- uniRankVar(varlistDe,
                  paste(outcome,"~1"),
    	            outcome,
                  DEdataframe,
                  rankingTest = "CStat")
    
    univarDeTest <- uniRankVar(varlistDe,
                  paste(outcome,"~1"),
    	            outcome,
                  DEdataframe,
                  testData=predTestDe,
                  rankingTest = "CStat")
    
    topfeat <- c(1:nrow(univarDe$orderframe)) <= topFeat
    posdisc <- (univarDe$orderframe$ROCAUC >= RAWaucTHR) & topfeat
    postestrain <- posdisc & (univarDeTest$orderframe$ROCAUC >= RAWaucTHR)
    allposIDeA <- c(allposIDeA,sum(postestrain)/(0.00001+sum(posdisc)))

    trainBACC <- (univarDe$orderframe$Sensitivity + univarDe$orderframe$Specificity)/2.0
    testBACC <- (univarDeTest$orderframe$Sensitivity + univarDeTest$orderframe$Specificity)/2.0
    ChangeIDeAAUCs <- c(ChangeIDeAAUCs,univarDeTest$orderframe$ROCAUC[posdisc] - univarDe$orderframe$ROCAUC[posdisc])
    ChangeIDeABACs <- c(ChangeIDeABACs,testBACC[posdisc] - trainBACC[posdisc])
    
    notinRaw <- (!(univarDe$orderframe$Name %in% colnames(dataframe)) & posdisc) 
    
    ChangeIDeAAUCs_LV <- c(ChangeIDeAAUCs_LV,univarDeTest$orderframe$ROCAUC[notinRaw] - univarDe$orderframe$ROCAUC[notinRaw])
    ChangeIDeABACs_LV <- c(ChangeIDeABACs_LV,testBACC[notinRaw] - trainBACC[notinRaw])
    
    posdisc_LV <- posdisc & notinRaw
    postestrain_LV <- posdisc_LV & (univarDeTest$orderframe$ROCAUC >= RAWaucTHR)
    allposIDeA_LV <- c(allposIDeA_LV,sum(postestrain_LV)/(0.00001+sum(posdisc_LV)))

    
    IDeAAUCs <- c(IDeAAUCs,univarDeTest$orderframe$ROCAUC[posdisc])
    IDeABACs <- c(IDeABACs,testBACC[posdisc])

    IDeAAUCs_LV <- c(IDeAAUCs_LV,univarDeTest$orderframe$ROCAUC[notinRaw])
    IDeABACs_LV <- c(IDeABACs_LV,testBACC[notinRaw])


    ####################### FCA
    
    iscontinous <- sapply(apply(trainDataFrame,2,unique),length) > 5 ## Only variables with enough samples
    
    noclassData <- trainDataFrame[,iscontinous]
    cmat <- cor(noclassData);
    diag(cmat) <- 0;
    maxcor <- apply(cmat,2,max);
    topcor <- names(maxcor[maxcor>rhoThreshold]) ## Only correlated features will be PCA
    cmat <- NULL
    
    notcorr <- colnames(trainDataFrame)[!(colnames(trainDataFrame) %in% topcor)]
    noclassData <- noclassData[,topcor]
    noclassTestData <- testDataFrame[,topcor]
    
    
    ## EFA
    
    uls <- fa(noclassData,nEFAFactors,rotate="varimax")  #unweighted least squares is minres 
    EFA_Train <- as.data.frame(cbind(predict(uls,noclassData),trainDataFrame[,notcorr]))
    colnames(EFA_Train) <- c(colnames(predict(uls,noclassData)),notcorr)
    EFA_Predicted <- as.data.frame(cbind(predict(uls,noclassTestData),testDataFrame[,notcorr]))
    colnames(EFA_Predicted) <- c(colnames(predict(uls,noclassTestData)),notcorr)
    
    iscontinous <- colnames(EFA_Predicted)[sapply(apply(EFA_Predicted,2,unique),length) > 5] ## Only variables with enough 
    varlistEFA <-  iscontinous
    univarEFA <- uniRankVar(varlistEFA,
                  paste(outcome,"~1"),
    	            outcome,
                  EFA_Train,
                  rankingTest = "CStat")
    
    univarEFATest <- uniRankVar(varlistEFA,
                  paste(outcome,"~1"),
    	            outcome,
                  EFA_Train,
                  testData=EFA_Predicted,
                  rankingTest = "CStat")
    
    topfeat <- c(1:nrow(univarEFA$orderframe)) <= topFeat
    posdisc <- (univarEFA$orderframe$ROCAUC >= RAWaucTHR) & topfeat
    postestrain <- posdisc & (univarEFATest$orderframe$ROCAUC >= RAWaucTHR)
    allposEFA <- c(allposEFA,sum(postestrain)/(0.00001+sum(posdisc)))

    trainBACC <- (univarEFA$orderframe$Sensitivity + univarEFA$orderframe$Specificity)/2.0
    testBACC <- (univarEFATest$orderframe$Sensitivity + univarEFATest$orderframe$Specificity)/2.0
    ChangeEFAAUCs <- c(ChangeEFAAUCs,univarEFATest$orderframe$ROCAUC[posdisc] - univarEFA$orderframe$ROCAUC[posdisc])
    ChangeEFABACs <- c(ChangeEFABACs,testBACC[posdisc] - trainBACC[posdisc])
    
    notinRaw <- (!(univarEFA$orderframe$Name %in% colnames(dataframe)) & posdisc) 
    
    ChangeEFAAUCs_LV <- c(ChangeEFAAUCs_LV,univarEFATest$orderframe$ROCAUC[notinRaw] - univarEFA$orderframe$ROCAUC[notinRaw])
    ChangeEFABACs_LV <- c(ChangeEFABACs_LV,testBACC[notinRaw] - trainBACC[notinRaw])

    posdisc_LV <- posdisc & notinRaw
    postestrain_LV <- posdisc_LV & (univarEFATest$orderframe$ROCAUC >= RAWaucTHR)
    allposEFA_LV <- c(allposEFA_LV,sum(postestrain_LV)/(0.00001+sum(posdisc_LV)))
    
    EFAAUCs <- c(EFAAUCs,univarEFATest$orderframe$ROCAUC[posdisc])
    EFABACs <- c(EFABACs,testBACC[posdisc])

    EFAAUCs_LV <- c(EFAAUCs_LV,univarEFATest$orderframe$ROCAUC[notinRaw])
    EFABACs_LV <- c(EFABACs_LV,testBACC[notinRaw])


}


```

```{r results = "asis"}

pander::pander(AUCthrlist)
pander::pander(mean(AUCthrlist))
```


### Violin of ROC AUC 
```{r results = "asis", warning = FALSE, dpi=300, fig.height= 4, fig.width= 5}

par(op)
par(mfrow=c(1,1),cex=0.7)

AUCs <- list(RAW=RawAUCs,IDeA=IDeAAUCs,IDeA_LV=IDeAAUCs_LV,EFA=EFAAUCs,EFA_LV=EFAAUCs_LV)
vioplot(AUCs,
        ylim=c(0.5,1.0),
        ylab="Test AUC",
        main="Test ROC AUC",
        col=colors,
        cex.axis=0.8
)
#stripchart(diffAUC, method = "jitter", col = "gray",
#           vertical = TRUE, pch = 1, add = TRUE,cex=0.35)
points(1:length(AUCs),lapply(AUCs,mean), pch = 18, col = "yellow", cex = 2.25)
par(op)


bp<- boxplot(AUCs,plot=FALSE)
bpstats <- as.data.frame(bp$stats)
rownames(bpstats) <- c("lower whisker","lower hinge","median","upper hinge","upper whisker")
colnames(bpstats) <- bp$names
pander::pander(bpstats)

```

### Violin of Balanced Accuracy 
```{r results = "asis", warning = FALSE, dpi=300, fig.height= 4, fig.width= 5}

par(op)
par(mfrow=c(1,1),cex=0.7)

BACs <- list(RAW=RawBACs,IDeA=IDeABACs,IDeA_LV=IDeABACs_LV,EFA=EFABACs,EFA_LV=EFABACs_LV)
vioplot(BACs,
        ylim=c(0.5,1.0),
        ylab="Test BAC",
        main="Test Balanced Accuracy",
        col=colors,
        cex.axis=0.8
)
#stripchart(diffBAC, method = "jitter", col = "gray",
#           vertical = TRUE, pch = 1, add = TRUE,cex=0.35)
points(1:length(BACs),lapply(BACs,mean), pch = 18, col = "yellow", cex = 2.25)
par(op)


bp<- boxplot(BACs,plot=FALSE)
bpstats <- as.data.frame(bp$stats)
rownames(bpstats) <- c("lower whisker","lower hinge","median","upper hinge","upper whisker")
colnames(bpstats) <- bp$names
pander::pander(bpstats)

```


### Violin of ROC AUC Changes
```{r results = "asis", warning = FALSE, dpi=300, fig.height= 4, fig.width= 5}

par(op)
colors=c("red","blue","darkblue","green","darkgreen")
par(mfrow=c(1,1),cex=0.7)

diffAUC <- list(RAW=ChangeRawAUCs,IDeA=ChangeIDeAAUCs,IDeA_LV=ChangeIDeAAUCs_LV,EFA=ChangeEFAAUCs,EFA_LV=ChangeEFAAUCs_LV)
vioplot(diffAUC,
        ylim=c(-0.5,0.5),
        ylab="Test-Train",
        main="Test-Train Paired ROC AUC",
        col=colors,
        cex.axis=0.8
)
#stripchart(diffAUC, method = "jitter", col = "gray",
#           vertical = TRUE, pch = 1, add = TRUE,cex=0.35)
points(1:length(diffAUC),lapply(diffAUC,mean), pch = 18, col = "yellow", cex = 2.25)
par(op)


bp<- boxplot(diffAUC,plot=FALSE)
bpstats <- as.data.frame(bp$stats)
rownames(bpstats) <- c("lower whisker","lower hinge","median","upper hinge","upper whisker")
colnames(bpstats) <- bp$names
pander::pander(bpstats)

```

### Violin of Balance Accuracy Changes
```{r results = "asis", warning = FALSE, dpi=300, fig.height= 4, fig.width= 5}


par(op)
par(mfrow=c(1,1),cex=0.7)

diffBAC <- list(RAW=ChangeRawBACs,IDeA=ChangeIDeABACs,IDeA_LV=ChangeIDeABACs_LV,EFA=ChangeEFABACs,EFA_LV=ChangeEFABACs_LV)
vioplot(diffBAC,
        ylim=c(-0.5,0.5),
        ylab="Test-Train",
        main="Test-Train Paired Balance Accuracy",
        col=colors,
        cex.axis=0.8
)
#stripchart(diffAUC, method = "jitter", col = "gray",
#           vertical = TRUE, pch = 1, add = TRUE,cex=0.35)
points(1:length(diffBAC),lapply(diffBAC,mean), pch = 18, col = "yellow", cex = 2.25)
par(op)

bp<- boxplot(diffBAC,plot=FALSE)
bpstats <- as.data.frame(bp$stats)
rownames(bpstats) <- c("lower whisker","lower hinge","median","upper hinge","upper whisker")
colnames(bpstats) <- bp$names
pander::pander(bpstats)


```


### Violin of Booth Positive
```{r results = "asis", warning = FALSE, dpi=300, fig.height= 4, fig.width= 5}


par(op)
par(mfrow=c(1,1),cex=0.7)

posFind <- list(RAW=allposRaw,IDeA=allposIDeA,IDeA_LV=allposIDeA_LV,EFA=allposEFA,EFA_LV=allposEFA_LV)
vioplot(posFind,
        ylim=c(0,1.0),
        ylab="TPR",
        main="Positive Discovery Rate",
        col=colors,
        cex.axis=0.8
)
stripchart(posFind, method = "jitter", col = "gray",
           vertical = TRUE, pch = 1, add = TRUE,cex=0.5)
points(1:length(posFind),lapply(posFind,mean), pch = 18, col = "yellow", cex = 2.25)
par(op)


bp<- boxplot(posFind,plot=FALSE)
bpstats <- as.data.frame(bp$stats)
rownames(bpstats) <- c("lower whisker","lower hinge","median","upper hinge","upper whisker")
colnames(bpstats) <- bp$names
pander::pander(bpstats)

```

### Violin of Booth Positive top 
```{r results = "asis", warning = FALSE, dpi=300, fig.height= 4, fig.width= 5}


par(op)
par(mfrow=c(1,1),cex=0.7)

posFind <- list(RAW=allposRaw,IDeA=allposIDeA,EFA=allposEFA)
vioplot(posFind,
        ylim=c(0,1.0),
        ylab="TPR",
        main="Positive Discovery Rate",
        col=colors[c(1,2,4)],
        cex.axis=0.8
)
stripchart(posFind, method = "jitter", col = "gray",
           vertical = TRUE, pch = 1, add = TRUE,cex=0.5)
points(1:length(posFind),lapply(posFind,mean), pch = 18, col = "yellow", cex = 2.25)
par(op)


bp<- boxplot(posFind,plot=FALSE)
bpstats <- as.data.frame(bp$stats)
rownames(bpstats) <- c("lower whisker","lower hinge","median","upper hinge","upper whisker")
colnames(bpstats) <- bp$names
pander::pander(bpstats)

```


