---
title: "Univariate: Saliva COVID-19"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
  word_document: 
    reference_docx: WordStyle_FRESA.docx
    toc: yes
    fig_caption: yes
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message = FALSE,comment = "#>")
```

# COVID_19 Univariate

### Loading the libraries

```{r}
library("FRESA.CAD")
library(readxl)
op <- par(no.readonly = TRUE)
pander::panderOptions('digits', 3)
pander::panderOptions('table.split.table', 400)
pander::panderOptions('keep.trailing.zeros',TRUE)

```


## Data: The COVID_19 Data-Set

The data to process is described in:

<https://zenodo.org/record/4156647#.Y1bSF3bMKUk>


Thermal Saliva Testing Dataset


10.5281/zenodo.4156647
<https://doi.org/10.5281/zenodo.4156647>


I added a column to the data identifying the repeated experiments.

```{r}

SalivaThermal <- as.data.frame(read_excel("~/GitHub/FCA/Data/SalivaThermal_Source_Data_2.xlsx"))


SalivaThermal_set1 <- subset(SalivaThermal,RepID==1)
rownames(SalivaThermal_set1) <- SalivaThermal_set1$ID
SalivaThermal_set1$RepID <- NULL
SalivaThermal_set1$ID <- NULL
SalivaThermal_set1$Ct <- NULL

SalivaThermal_set2 <- subset(SalivaThermal,RepID==2)
rownames(SalivaThermal_set2) <- SalivaThermal_set2$ID
SalivaThermal_set2$RepID <- NULL
SalivaThermal_set2$ID <- NULL
SalivaThermal_set2$Ct <- NULL

SalivaThermal_set3 <- subset(SalivaThermal,RepID==3)
rownames(SalivaThermal_set3) <- SalivaThermal_set3$ID
SalivaThermal_set3$RepID <- NULL
SalivaThermal_set3$ID <- NULL
SalivaThermal_set3$Ct <- NULL

SalivaThermal_Avg <- (SalivaThermal_set1 + SalivaThermal_set2 + SalivaThermal_set3)/3

colnames(SalivaThermal_Avg) <- paste("V",colnames(SalivaThermal_Avg),sep="_")

SalivaThermal_Avg$class <- 1*(str_detect(rownames(SalivaThermal_Avg),"P"))

```

#### Standarize the names for the reporting

```{r results = "asis"}
dataframe <- SalivaThermal_Avg
outcome <- "class"

```



```{r results = "asis"}
pander::pander(c(rows=nrow(dataframe),col=ncol(dataframe)-1))
pander::pander(table(dataframe[,outcome]))

varlist <- colnames(dataframe)
varlist <- varlist[varlist != outcome]
varlist <- as.data.frame(cbind(name=varlist,desc=varlist))

```


## Univariate

```{r results = "asis"}

univariate_columns <- c("caseMean","caseStd","controlMean","controlStd","controlKSP","ROCAUC","WilcoxRes.p","FRes.p")
univar <- uniRankVar(varlist,
	           paste(outcome,"~1"),
	           outcome,
	           dataframe,
	           categorizationType = "Raw",
	           type = "LOGIT",
	           rankingTest = "AUC",
	            cateGroups = c(0.1, 0.9),
	            raw.dataFrame = NULL,
	            description = ".",
	            uniType = "Binary")

pander::pander(univar$orderframe[1:20,univariate_columns])

topfiveOrg <- rownames(univar$orderframe[1:5,])

```

## Decorrelation Analysis

```{r results = "asis"}
DEdataframe <- IDeA(dataframe,thr=0.80,verbose = TRUE)

pander::pander(c(Decorrleated_Fraction=sum(str_detect(colnames(DEdataframe),"La_"))/(ncol(DEdataframe)-1)))
demat <- attr(DEdataframe,"UPSTM")
pander::pander(c(sparse_Fraction=1.0-sum(demat==0)/ncol(demat)/nrow(demat)))
varlistDe <-  colnames(DEdataframe)[colnames(DEdataframe) != "class"];
varlistDe <- as.data.frame(cbind(name=varlistDe,desc=varlistDe))



```



### The heat maps.

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}

par(cex=0.6,cex.main=0.85,cex.axis=0.7)
cormat <- cor(dataframe[,rownames(demat)],method="spearman")
cormat[is.na(cormat)] <- 0
gplots::heatmap.2(abs(cormat),
                  trace = "none",
#                  scale = "row",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  main = "Spearman Correlation Original",
                  cexRow = 0.35,
                  cexCol = 0.35,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")


cormat <- cor(DEdataframe[,colnames(demat)],method="spearman")
cormat[is.na(cormat)] <- 0
gplots::heatmap.2(abs(cormat),
                  trace = "none",
#                  scale = "none",
                  mar = c(5,5),
                  col=rev(heat.colors(5)),
                  main = "Spearman Correlation: After UPSTM",
                  cexRow = 0.35,
                  cexCol = 0.35,
                  key.title=NA,
                  key.xlab="Spearman Correlation",
                  xlab="Feature", ylab="Feature")

par(op)
```


## Univariate Decorrelated

```{r results = "asis"}

univarDe <- uniRankVar(varlistDe,
              paste(outcome,"~1"),
	            outcome,
              DEdataframe,
	            categorizationType = "Raw",
	            type = "LOGIT",
	            rankingTest = "AUC",
	            cateGroups = c(0.1, 0.9),
	            raw.dataFrame = NULL,
	            description = ".",
	            uniType = "Binary")

pander::pander(univarDe$orderframe[1:20,univariate_columns])

```

### Comparing Decorrelation vs Original

```{r results = "asis"}
pthr <- 0.20/(ncol(dataframe)-1)

topDecorNames <- rownames(univarDe$orderframe[univarDe$orderframe$FRes.p<pthr,])
topDecorNames <- unique(c(topDecorNames,rownames(univarDe$orderframe[1:5,])))

#topDecorNames <- rownames(univarDe$orderframe[univarDe$orderframe$FRes.p<1.0e-5,])
dc <- getLatentCoefficients(DEdataframe)
### 2a Get only the ones that in the top features
deNames_in_dc <- topDecorNames[topDecorNames %in% names(dc)]
selectedlist <- dc[deNames_in_dc]
theDeFormulas <- selectedlist

rawuniv <- univariate_Wilcoxon(dataframe,outcome,limit=-1)
deuniv <- univariate_Wilcoxon(DEdataframe,outcome,limit=-1)
matsize <- (ncol(dataframe)-1)^2
nocorrelated <- ncol(dataframe) - 1 - ncol(demat)
pander::pander(c(sparse_Fraction=(sum(demat != 0) + nocorrelated)/matsize))
pander::pander(c(raw=length(rawuniv),decor=length(deuniv)))
pander::pander(c(Number_Latent=length(dc)))
pander::pander(c(meanSize=mean(sapply(dc,length))))

```


## CV ROC Analysis

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 4.0, fig.width= 8.0}
par(op)
par(mfrow=c(1,2),cex=0.9)
fraction <-0.70
repetitions <- 100

fcout <- round(fraction*nrow(dataframe)/15+1.0)
pander::pander(c(NumberofFeatures=fcout))

cvRaw <- randomCV(dataframe,
                outcome,
                fittingFunction= filteredFit,
                classSamplingType = "Pro",
                trainFraction = fraction,
                repetitions = repetitions,
#                fitmethod=  glm,
                fitmethod=  KNN_method,
                filtermethod=mRMR.classic_FRESA,
                filtermethod.control=list(feature_count= fcout),
#                family="binomial"
)
bpraw <- predictionStats_binary(cvRaw$medianTest,"RAW",cex=0.75)


pander::pander(bpraw$CM.analysis$tab)
pander::pander(bpraw$accc)
pander::pander(bpraw$aucs)
pander::pander(bpraw$berror)

cvDe <- randomCV(DEdataframe,
                outcome,
                fittingFunction= filteredFit,
                trainSampleSets= cvRaw$trainSamplesSets,
#                fitmethod=  glm,
                fitmethod=  KNN_method,
                filtermethod=mRMR.classic_FRESA,
                filtermethod.control=list(feature_count= fcout),
#                family="binomial"
)
bpDecor <- predictionStats_binary(cvDe$medianTest,"Decorrelated",cex=0.75)

par(op)

pander::pander(bpDecor$CM.analysis$tab)
pander::pander(bpDecor$accc)
pander::pander(bpDecor$aucs)
pander::pander(bpDecor$berror)

### Here we compute the probability that the outcome-driven decorrelation ROC is superior to the RAW ROC. 
pander::pander(roc.test(bpDecor$ROC.analysis$roc.predictor,bpraw$ROC.analysis$roc.predictor,alternative = "greater"))


```

## Feature Frequency Plots
```{r results = "asis", warning = FALSE, dpi=300, fig.height= 6.0, fig.width= 6.0}
par(mfrow=c(2,1),cex=0.9,cex.axis=0.8)

rawtopf <- cvRaw$featureFrequency/repetitions
crawtopf <- rawtopf

if (length(rawtopf) > 30)
{
  rawtopf <- rawtopf[1:30]
}
barplot(100*rawtopf,las=2,main="Raw Features",ylim=c(0,100.0),cex.names = 0.35,cex.axis = 0.35,ylab="Sel. %")

detopf <- cvDe$featureFrequency/repetitions
cdetopf <- detopf
names(cdetopf) <- str_remove_all(names(cdetopf),"Ba_")
names(cdetopf) <- str_remove_all(names(cdetopf),"La_")
if (length(detopf) > 30)
{
  detopf <- detopf[1:30]
}

barplot(100*detopf,las=2,main="Decorrelated Features",ylim=c(0,100.0),cex.names = 0.35,cex.axis = 0.35,ylab="Sel. %")

par(op)



```



### Final Table

```{r results = "asis"}
unlistdecorr <- selectedlist
names(unlistdecorr) <- NULL
unlistdecorr <- unique(names(unlist(unlistdecorr)))

finalTableDe <- univarDe$orderframe[deNames_in_dc,univariate_columns]

finalTableOr <- univar$orderframe[unique(c(topfiveOrg,unlistdecorr,names(crawtopf)[1:2],names(cdetopf)[1:2])),univariate_columns]

finalTable <- rbind(finalTableOr,finalTableDe)

deFromula <- character(length(theDeFormulas))
names(deFromula) <- names(theDeFormulas)

for (dx in names(deFromula))
{
  coef <- theDeFormulas[[dx]]
  cname <- names(theDeFormulas[[dx]])
  names(cname) <- cname
  for (cf in names(coef))
  {
    if (cf != dx)
    {
      if (coef[cf]>0)
      {
        deFromula[dx] <- paste(deFromula[dx],
                               sprintf("+ %5.3f*%s",coef[cf],cname[cf]))
      }
      else
      {
        deFromula[dx] <- paste(deFromula[dx],
                               sprintf("%5.3f*%s",coef[cf],cname[cf]))
      }
    }
  }
}

orgnamez <- rownames(finalTable)
orgnamez <- str_remove_all(orgnamez,"Ba_")
orgnamez <- str_remove_all(orgnamez,"La_")
finalTable$uAUC <- univar$orderframe[orgnamez,"ROCAUC"]
finalTable$raw_Freq <- crawtopf[orgnamez]
finalTable$La_Freq <- cdetopf[orgnamez]
finalTable$DecorFormula <- deFromula[rownames(finalTable)]
fscores <- attr(DEdataframe,"fscore")
finalTable$fscores <- fscores[rownames(finalTable)]

finalTable <- finalTable[order(-finalTable$ROCAUC),]

pander::pander(finalTable)
```


