---
title: "ILAA Tutorial: Body_fat"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
  word_document: 
    reference_docx: WordStyle_FRESA.docx
    toc: yes
    fig_caption: yes
editor_options:
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message = FALSE,comment = "#>")

pander::panderOptions('digits', 3)
pander::panderOptions('keep.trailing.zeros',TRUE)
op <- par(no.readonly = TRUE)

```

# ILAA Tutorial

Iterative Linear Association Analysis (ILAA) is a computational method
that creates a linear transformation from tabular data that effectively
removes linear associations between the data-set variables. The returned
transformation matrix can be used to:

1.  Exploratory analysis of latent variables and their association to
    all the observed variables

2.  Exploratory discovery of latent variables associated with an
    specific outcome-target

3.  Address multicollarity issues in linear regression models:

    1.  Better estimation and interpretation of model variables

    2.  Improve linear model performance

4.  Simplify the multidimensional search space aiding the creation of
    complex ML models

The objective of this tutorial is to guide users in the use of the ILAA
to effectively accomplish the above mentioned tasks. The tutorial will
showcase:

-   Transform a data-frame affected by data multicollinearity into a new
    data-frame with minimum data correlation among variables

-   Visualize the transformation matrix

-   Explore the returned formulas for each one of the returned latent
    variables

-   Understand and interpret the returned latent variables

-   Use ILAA as preprocessing step to model an specific target-outcome
    using linear models

    -   Explore the model in transformed space

    -   Get the corresponding model based on the observed variables.

## The Libraries

ILAA is a wrapper of the more general method of data decorrelation
(IDeA) implemented in R and both are part of the FRESA.CAD 3.4.6
package.

```{r}
## From git hub
#install_github("joseTamezPena/FRESA.CAD")

library("FRESA.CAD")

```

## Material and Methods

For this tutorial I'll use the body-fat prediction data set. The data
was downloaded from Kaggle:

<https://www.kaggle.com/datasets/fedesoriano/body-fat-prediction-dataset>

I'm coping the kaggle data disclaimer:

"Source The data were generously supplied by Dr. A. Garth Fisher who
gave permission to freely distribute the data and use for non-commercial
purposes.

Roger W. Johnson Department of Mathematics & Computer Science South
Dakota School of Mines & Technology 501 East St. Joseph Street Rapid
City, SD 57701

email address:
[rwjohnso\@silver.sdsmt.edu](mailto:rwjohnso@silver.sdsmt.edu){.email}
web address: <http://silver.sdsmt.edu/~rwjohnso>"

## Loading the Data

The following code loads the data and removes the density information
from the data. It aslo computes the Body Mass Index (BMI)

```{r}
body_fat <- read.csv("~/GitHub/LatentBiomarkers/Data/BodyFat/BodyFat.csv", header=TRUE)

### Removing density as estimator
body_fat$Density <- NULL

body_fat$BMI <- 10000*body_fat$Weight*0.453592/((body_fat$Height*2.54)^2)
## Removing subjects with data errors
body_fat <- body_fat[body_fat$BMI<=50,]

```

## ILAA unsupervised processing

The ILAA function is:

``` r
 ILAA(data=NULL,
                thr=0.80,
                method=c("pearson","spearman"),
                Outcome=NULL,
                drivingFeatures=NULL,
                maxLoops=100,
                verbose=FALSE,
                ...)
```

where:

-   `data`: The source data-frame

-   *`thr`* : The target correlation goal.

-   *`method`* : Defines the correlation measure

-   `Outcome` The name of the target variable, and it is required for
    supervised learning

-   `drivingFeatures` : Defines a set of variables that are aimed to be
    basis unaltered vectors

-   `maxLoops` : The maximum number of iterations cycles

-   `verbose` : Display the evolution of the algorithm.

-   `…` : Parameters to be passed to the `featureAdjustment()` method

By default the ILAA function will target a correlation lower to 0.8
using the pearson correlation measure.

For this example, I'll set the correlation goal to 0.2 in verbose mode:

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}

# Calling ILAA to achieve a final correlation of 0.2
body_fat_Decorrelated <- ILAA(body_fat,thr=0.2,verbose=TRUE)

```

### Data Frame Attributes

``` r
  attr(dataTransformed,"UPSTM") : The transformation matrix
  attr(dataTransformed,"fscore") : The score of each feature
  attr(dataTransformed,"drivingFeatures") : The list of driving features
  attr(dataTransformed,"unaltered") : The list of unaltered features
  attr(dataTransformed,"LatentVariables") : The list of latent variables
  attr(dataTransformed,"R.critical") : The estimated minium correlation
  attr(dataTransformed,"IDeAEvolution") : Evolution of the algorithm
```

The main attributes is "`UPSTM`". That stores the transformation matrix
from observed variables to the latent variable. The "`IDeAEvolution`"
attribute can be used to verify if the algorithm achieved the target
correlation goal, and the sparsity of the returned matrix.

### Plotting the Evolution

Here we will use the `attr(dataTransformed,"IDeAEvolution")` to plot the
evolution of the correlation measure and the evolution of the matrix
sparsity.

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 3.5, fig.width= 7.0}
par(mfrow=c(1,2),cex=0.5)

# Correlation
yval <- attr(body_fat_Decorrelated,"IDeAEvolution")$Corr
xidx <- c(1:length(yval))
plot(xidx,yval,
     xlab="Iteration Cycle",
     ylab="Max. Pearson Correlation",
     ylim=c(0,1.0),
     main="Evolution of the maximum Correlation")
  lfit <-try(loess(yval~xidx,span=0.5));
  if (!inherits(lfit,"try-error"))
  {
    plx <- try(predict(lfit,se=TRUE))
    if (!inherits(plx,"try-error"))
    {
      lines(xidx,plx$fit,lty=1,col="red")
    }
  }

# Sparsity  
yval <- attr(body_fat_Decorrelated,"IDeAEvolution")$Spar

plot(xidx,yval,
     xlab="Iteration Cycle",
     ylab="Matrix Sparcity",
     ylim=c(0,1.0),
     main="Evolution of the Matrix Sparcity")
  lfit <-try(loess(yval~xidx,span=0.5));
  if (!inherits(lfit,"try-error"))
  {
    plx <- try(predict(lfit,se=TRUE))
    if (!inherits(plx,"try-error"))
    {
      lines(xidx,plx$fit,lty=1,col="red")
    }
  }

```

### The ILAA transformed data

Before exploring into more detail the properties of the `ILAA` results.
Let us first verify that the returned matrix does not contain features
with very high correlation among them.

Here I'll plot the original correlation and the correlation of the
returned data set.

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}

# The original
  par(cex=0.6,cex.main=0.85,cex.axis=0.7)
  cormat <- cor(body_fat,method="pearson")
  gplots::heatmap.2(abs(cormat),
                    trace = "none",
                    mar = c(5,5),
                    col=rev(heat.colors(5)),
                    main = "Original Correlation",
                    cexRow = 0.75,
                    cexCol = 0.75,
                     srtCol=30,
                     srtRow=60,
                    key.title=NA,
                    key.xlab="|Pearson Correlation|",
                    xlab="Feature", ylab="Feature")

# The transformed
  cormat <- cor(body_fat_Decorrelated,method="pearson")
  gplots::heatmap.2(abs(cormat),
                    trace = "none",
                    mar = c(5,5),
                    col=rev(heat.colors(5)),
                    main = "After ILAA Correlation",
                    cexRow = 0.75,
                    cexCol = 0.75,
                     srtCol=30,
                     srtRow=60,
                    key.title=NA,
                    key.xlab="|Pearson Correlation|",
                    xlab="Feature", ylab="Feature")

```

### Exploring transformation

The `attr(body_fat_Decorrelated,"UPSTM")` returns the transformation
matrix. The `UPSTM` is sparse, here I show a heat-map of the
transformation matrix that shows which elements are different from zero.

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}

  UPSTM <- attr(body_fat_Decorrelated,"UPSTM")
  
  gplots::heatmap.2(1.0*(abs(UPSTM)>0),
                    trace = "none",
                    mar = c(5,5),
                    col=rev(heat.colors(5)),
                    main = "Decorrelation matrix",
                    cexRow = 0.75,
                    cexCol = 0.75,
                   srtCol=30,
                   srtRow=60,
                    key.title=NA,
                    key.xlab="|Beta|>0",
                    xlab="Output Feature", ylab="Input Feature")
  
```

### The Latent Formulas

The sparsity of the `UPSTM` matrix can be analyzed to get the formula
for each on of the latent formulas. The `getLatentCoefficients()` and
its attribute: `attr(LatentFormulas,"LatentCharFormulas")` are used to
display the shape of the latent variables.

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}
# Get a list with the latent formulas' coefficients
LatentFormulas <- getLatentCoefficients(body_fat_Decorrelated)

# A string character with the formulas can be obtained by:
charFormulas <- attr(LatentFormulas,"LatentCharFormulas")
pander::pander(as.matrix(charFormulas))

```

### Latent Variable Interpretation

The ILAA returns the Unit Preserving Spatial Transformation Matrix
(UPSTM). This specific transformation is the combination of
statistically significant linear association analysis between feature
pairs. Each significant association is modeled by a linear equation;
henceforth, the interpretation of each feature is as follows:

-   Each discovered latent variable is the residual of the observed
    parent variable *vs.* the model of the variables associated with the
    parent variable. For example: $$
    Lawrist = Wrist + 0.017BodyFat - 0.026Weight.  
    $$

    Describes that the $Wrist$ is associated with the $BodyFat$ and the
    $Weight$, and the latent variable $Lawrist$ is the amount of
    information in the $Wrist$ not found by $BodyFat$ nor the $Weight$.

-   The model of the $Wrist$ in therefore:

$$
Wrist = -0.017BodyFat + 0.026Weight.  
$$

The following lines of code shows the association of the latent variable
to each one of the observed parent variable, and the association of the
parent variable to the linear model.

```{r warning = FALSE, dpi=300, fig.height= 3.5, fig.width= 7.0}

par(mfrow=c(1,2),cex=0.35)
fnames <- names(charFormulas)[1]
for (fnames in names(charFormulas))
{
  obsname <- str_remove(fnames,"La_")
  menv <- mean(body_fat_Decorrelated[,fnames])
  range <- max(body_fat[,obsname])-min(body_fat[,obsname])
  ylim <- c(menv-range/2,menv+range/2)
  plot(body_fat[,obsname],body_fat_Decorrelated[,fnames],
       ylim=ylim,
       ylab=fnames,
       xlab=obsname,
       main=paste("ILAA Latent Variable:",fnames))
  deformula <- LatentFormulas[[fnames]]
  noInames <- names(deformula)[names(deformula) != obsname]
  predObs <- -(as.matrix(body_fat[,noInames]) %*% deformula[noInames])
  plot(predObs,body_fat[,obsname],ylab=obsname,xlab=charFormulas[fnames],
       main=paste("ILAA Generated Predictions of",obsname),cex.labels=0.5)
}

par(op)

```

## ILAA for Supervised Learning

### Split into Training Testing Sets

```{r}

# 75% for training 25% for testing 
set.seed(1)
trainsamples <- sample(nrow(body_fat),3*nrow(body_fat)/4)

trainingset <- body_fat[trainsamples,]
testingset <- body_fat[-trainsamples,]

```

## Data decorrelation of the train set and prediction of the test set

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}

body_fat_Decorrelated_train <- ILAA(trainingset,thr=0.2,Outcome="BodyFat",drivingFeatures="NA")
body_fat_Decorrelated_test <- predictDecorrelate(body_fat_Decorrelated_train,testingset)

```

### Train a regression model for body fat prediction

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}

modelBodyFat <- LASSO_1SE(BodyFat~.,body_fat_Decorrelated_train)
pander::pander(as.matrix(modelBodyFat$coef))

```

### The model coefficients in the observed space

```{r  results = "asis", warning = FALSE}

# Get the coefficients in the observed space
observedCoef <- getObservedCoef(body_fat_Decorrelated_train,modelBodyFat)
pander::pander(as.matrix(observedCoef$coefficients))


```

### Predict using the transformed data set

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}

predicBodyFat <- predict(modelBodyFat,body_fat_Decorrelated_test)
rmetrics <- predictionStats_regression(cbind(testingset$BodyFat,predicBodyFat),"Body Fat")
pander::pander(rmetrics)

```

### Prediction using the observed features

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}


predicBodyFatObst <- model.matrix(formula(observedCoef$formula),testingset) %*% observedCoef$coefficients

plot(predicBodyFatObst,predicBodyFat,xlab="Observed",ylab="Latent",main="Observed vs Latent")


```

### Comparison to raw model

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}
rawmodelBodyFat <- LASSO_1SE(BodyFat~.,trainingset)
pander::pander(rawmodelBodyFat$coef)
rawpredicBodyFat <- predict(rawmodelBodyFat,testingset)
rmetrics <- predictionStats_regression(cbind(testingset$BodyFat,rawpredicBodyFat),"Body Fat")
pander::pander(rmetrics)

```

### Comparing the feature significance on the model

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}

par(mfrow=c(2,2),cex=0.5)

rawlm <- lm(BodyFat~.,trainingset[,c("BodyFat",names(rawmodelBodyFat$coef)[-1])])
pander::pander(rawlm,add.significance.stars=TRUE)
plot(rawlm)

par(mfrow=c(2,2),cex=0.5)
Delm <- lm(BodyFat~.,body_fat_Decorrelated_train[,c("BodyFat",names(modelBodyFat$coef)[-1])])
pander::pander(Delm,add.significance.stars=TRUE)
plot(Delm)

par(op)
```

## Train a regression model for Overweight prediction

### First remove Height and Weight from training and testing sets

```{r}

trainingsetBMI <- trainingset[,!(colnames(trainingset) %in% c("Weight","Height"))]
testingsetBMI <- testingset[,!(colnames(trainingset) %in% c("Weight","Height"))]
trainingsetBMI$Overweight <- 1*(trainingsetBMI$BMI>=25)
testingsetBMI$Overweight <- 1*(testingsetBMI$BMI>=25)
trainingsetBMI$BMI <- NULL
testingsetBMI$BMI <- NULL
table(trainingsetBMI$Overweight)
table(testingsetBMI$Overweight)

BMI_Decorrelated_train <- ILAA(trainingsetBMI,thr=0.2,Outcome="Overweight",drivingFeatures="NA")

BMI_Decorrelated_test <- predictDecorrelate(BMI_Decorrelated_train,testingsetBMI)

```

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}

modelOverweight <- LASSO_1SE(Overweight~.,BMI_Decorrelated_train,family="binomial")
pander::pander(as.matrix(modelOverweight$coef))


```

### The model coefficients in the observed space

```{r  results = "asis", warning = FALSE}

# Get the coefficients in the observed space
observedCoef <- getObservedCoef(BMI_Decorrelated_train,modelOverweight)
pander::pander(as.matrix(observedCoef$coefficients))


```

### Predict using the transformed data set

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}

predicOverweight <- predict(modelOverweight,BMI_Decorrelated_test)
pr <- predictionStats_binary(cbind(BMI_Decorrelated_test$Overweight,predicOverweight),"Overweight")
pander::pander(pr$ClassMetrics)

```

### Prediction using the observed features

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}

predicOverweightObst <- model.matrix(formula(observedCoef$formula),testingsetBMI) %*% observedCoef$coefficients
#predicOverweightObst <- 1.0/(1.0 + exp(-predicOverweightObst));

plot(predicOverweightObst,predicOverweight,xlab="Observed",ylab="Latent",main="Observed vs Latent")


```

### Comparison to raw model

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}
rawmodelOverweight <- LASSO_1SE(Overweight~.,trainingsetBMI,family="binomial")
pander::pander(rawmodelOverweight$coef)
rawpredicOverweight <- predict(rawmodelOverweight,testingsetBMI)
pr <- predictionStats_binary(cbind(testingsetBMI$Overweight,rawpredicOverweight),"Overweight")
pander::pander(pr$ClassMetrics)

```

### Comparing the feature significance on the model

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}

par(mfrow=c(2,2),cex=0.5)

rawlm <- lm(Overweight~.,trainingsetBMI[,c("Overweight",names(rawmodelOverweight$coef)[-1])])
pander::pander(rawlm,add.significance.stars=TRUE)
plot(rawlm)

par(mfrow=c(2,2),cex=0.5)
Delm <- lm(Overweight~.,BMI_Decorrelated_test[,c("Overweight",names(modelOverweight$coef)[-1])])
pander::pander(Delm,add.significance.stars=TRUE)
plot(Delm)

```
