---
title: "Decorrelation-Based Feature Discovery: body_fat"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
  word_document: 
    reference_docx: WordStyle_FRESA.docx
    toc: yes
    fig_caption: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message = FALSE,comment = "#>")

```


# Effect of UPSTM-Based Decorrelation on Feature Discovery


### Loading the libraries

```{r}
library("FRESA.CAD")
library(readxl)
library(igraph)
library(umap)
library(tsne)
library(entropy)
library(psych)
library(whitening)
library("vioplot")
library("rpart")

op <- par(no.readonly = TRUE)
pander::panderOptions('digits', 3)
pander::panderOptions('table.split.table', 400)
pander::panderOptions('keep.trailing.zeros',TRUE)

```

## Material and Methods

Data Source 
https://www.kaggle.com/datasets/fedesoriano/body-fat-prediction-dataset


"Source
The data were generously supplied by Dr. A. Garth Fisher who gave permission to freely distribute the data and use for non-commercial purposes.

Roger W. Johnson
Department of Mathematics & Computer Science
South Dakota School of Mines & Technology
501 East St. Joseph Street
Rapid City, SD 57701

email address: rwjohnso@silver.sdsmt.edu
web address: http://silver.sdsmt.edu/~rwjohnso"


## The Data

```{r}
body_fat <- read.csv("~/GitHub/LatentBiomarkers/Data/BodyFat/BodyFat.csv", header=TRUE)

### Removing density as estimator
body_fat$Density <- NULL

#par(cex=0.5)
```

#### Standarize the names for the reporting

```{r results = "asis"}
studyName <- "Body_Fat"
dataframe <- body_fat
outcome <- "BodyFat"

thro <- 0.35
cexheat <- 0.75
loops  <- 30
```

## Generaring the report

```{r parent, child = here::here("~/GitHub/LatentBiomarkers/RMD/Child_univariate/child_regression.Rmd")}
```


### The Body Fat plots

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}

plot(predOutcome,testingset[,outcome],xlab="Raw Predicted",ylab=outcome,main="Raw Fat Prediction")
plot(predOutcome_DE,testingset[,outcome],xlab="IDeA Predicted",ylab=outcome,main="IDeA Fat Prediction")

```


