---
title: "Decorrelation-Based Feature Discovery: Activity"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
  word_document: 
    reference_docx: WordStyle_FRESA.docx
    toc: yes
    fig_caption: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message = FALSE,comment = "#>")

```

# Effect of UPSTM-Based Decorrelation on Feature Discovery


### Loading the libraries

```{r}
library("FRESA.CAD")
library(readxl)
library(igraph)
library(umap)
library(tsne)
library(entropy)

op <- par(no.readonly = TRUE)
pander::panderOptions('digits', 3)
pander::panderOptions('table.split.table', 400)
pander::panderOptions('keep.trailing.zeros',TRUE)
source("C:/Users/jtame/Documents/GitHub/LatentBiomarkers/RMD/DecorrelationCV.R")

```

## Material and Methods

Data Source 
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6960825/

Mohino-Herranz I, Gil-Pita R, Rosa-Zurera M, Seoane F. Activity Recognition Using Wearable Physiological Measurements: Selection of Features from a Comprehensive Literature Study. Sensors (Basel). 2019 Dec 13;19(24):0. doi: 10.3390/s19245524. PMID: 31847261; PMCID: PMC6960825.



## The Data

```{r}
Activitydata <- read.csv("~/GitHub/LatentBiomarkers/Data/ActivityData/data.txt", header=FALSE, stringsAsFactors=TRUE)
featNames <- read.table("~/GitHub/LatentBiomarkers/Data/ActivityData/Featurelabels.txt", quote="\"", comment.char="")
featNames <- as.character(t(featNames))
featNames <- str_replace_all(featNames,"\\(abs\\)","_A_")
featNames[duplicated(featNames)] <- paste(featNames[duplicated(featNames)],"D",sep="_")

rep_ID <- numeric(nrow(Activitydata))
ctr <- 1
for (ind in c(1:(nrow(Activitydata)-1)))
{
  rep_ID[ind] <- ctr
  if (Activitydata$V1[ind] != Activitydata$V1[ind+1]) ctr <- 0;
  ctr <- ctr + 1
}
rownames(Activitydata) <- paste(Activitydata$V1,rep_ID,sep="_")
colnames(Activitydata) <- c("ID",featNames,"class")
Activitydata$rep <- rep_ID
  
tb <- table(Activitydata$class)

classes <- c("Neu","Emo","Men","Phy")
names(classes) <- names(tb)
Activitydata$class <- classes[as.character(Activitydata$class)]
table(Activitydata$class)



ID_class <- paste(Activitydata$ID,Activitydata$class)
IDCLASS <- unique(ID_class)
theclass <- Activitydata$class[!duplicated(ID_class)]
theIDs <- Activitydata$ID[!duplicated(ID_class)]

ActivitydataAvg <- NULL
for (id in IDCLASS)
{
  ActivitydataAvg <- rbind(ActivitydataAvg,apply(Activitydata[ID_class==id,featNames],2,mean))
}
colnames(ActivitydataAvg) <- featNames
rownames(ActivitydataAvg) <- IDCLASS
ActivitydataAvg <- as.data.frame(ActivitydataAvg)
ActivitydataAvg$class <- theclass
ActivitydataAvg$ID <- theIDs

table(ActivitydataAvg$class)


ActivitydataAvg <- subset(ActivitydataAvg, class=="Men" | class=="Emo")

ActivitydataAvg$class <- 1*(ActivitydataAvg$class == "Men")
table(ActivitydataAvg$class)

```


#### Standarize the names for the reporting

```{r results = "asis"}
studyName <- "Activity"
dataframe <- ActivitydataAvg
outcome <- "class"
nLoops <- 100

needScaling <- TRUE
IDSample <- TRUE
thro <- 0.8

```
## CV Analysis and report

```{r parent, child = here::here("~/GitHub/LatentBiomarkers/RMD/Child_univariate/CV_analysis.Rmd")}
```


## Testing different parameters
```{r}
nLoops <- 30

cvData_0.9 <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.9)
cvData_0.8 <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.8)
cvData_0.7 <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.7)
cvData_0.6 <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.6)
cvData_0.5 <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.5)
bpDe <- predictionStats_binary(cvData_0.9$DeMedTestRaw,"IDeA@0.9")
bpDe <- predictionStats_binary(cvData_0.8$DeMedTestRaw,"IDeA@0.8")
bpDe <- predictionStats_binary(cvData_0.7$DeMedTestRaw,"IDeA@0.7")
bpDe <- predictionStats_binary(cvData_0.6$DeMedTestRaw,"IDeA@0.6")
bpDe <- predictionStats_binary(cvData_0.5$DeMedTestRaw,"IDeA@0.5")

cvData_0.9_R <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.9,corRank=FALSE)
cvData_0.8_R <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.8,corRank=FALSE)
cvData_0.7_R <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.7,corRank=FALSE)
cvData_0.6_R <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.6,corRank=FALSE)
cvData_0.5_R <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.5,corRank=FALSE)
bpDe <- predictionStats_binary(cvData_0.9_R$DeMedTestRaw,"IDeA@0.9 CorRank")
bpDe <- predictionStats_binary(cvData_0.8_R$DeMedTestRaw,"IDeA@0.8 CorRank")
bpDe <- predictionStats_binary(cvData_0.7_R$DeMedTestRaw,"IDeA@0.7 CorRank")
bpDe <- predictionStats_binary(cvData_0.6_R$DeMedTestRaw,"IDeA@0.6 CorRank")
bpDe <- predictionStats_binary(cvData_0.5_R$DeMedTestRaw,"IDeA@0.5 CorRank")


cvData_0.9_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                       DecorPCAEFA=FALSE,thr=0.9,Outcome=outcome,drivingFeatures=outcome)
cvData_0.8_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                       DecorPCAEFA=FALSE,thr=0.8,Outcome=outcome,drivingFeatures=outcome)
cvData_0.7_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                       DecorPCAEFA=FALSE,thr=0.7,Outcome=outcome,drivingFeatures=outcome)
cvData_0.6_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                       DecorPCAEFA=FALSE,thr=0.6,Outcome=outcome,drivingFeatures=outcome)
cvData_0.5_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                       DecorPCAEFA=FALSE,thr=0.5,Outcome=outcome,drivingFeatures=outcome)
bpDe <- predictionStats_binary(cvData_0.9_D$DeMedTestRaw,"IDeA@0.9 Driven")
bpDe <- predictionStats_binary(cvData_0.8_D$DeMedTestRaw,"IDeA@0.8 Driven")
bpDe <- predictionStats_binary(cvData_0.7_D$DeMedTestRaw,"IDeA@0.7 Driven")
bpDe <- predictionStats_binary(cvData_0.6_D$DeMedTestRaw,"IDeA@0.6 Driven")
bpDe <- predictionStats_binary(cvData_0.5_D$DeMedTestRaw,"IDeA@0.5 Driven")

cvData_0.9_S <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.9,method="spearman")
cvData_0.8_S <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.8,method="spearman")
cvData_0.7_S <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.7,method="spearman")
cvData_0.6_S <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.6,method="spearman")
cvData_0.5_S <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.5,method="spearman")
bpDe <- predictionStats_binary(cvData_0.9_S$DeMedTestRaw,"IDeA@0.9 Spearman")
bpDe <- predictionStats_binary(cvData_0.8_S$DeMedTestRaw,"IDeA@0.8 Spearman")
bpDe <- predictionStats_binary(cvData_0.7_S$DeMedTestRaw,"IDeA@0.7 Spearman")
bpDe <- predictionStats_binary(cvData_0.6_S$DeMedTestRaw,"IDeA@0.6 Spearman")
bpDe <- predictionStats_binary(cvData_0.5_S$DeMedTestRaw,"IDeA@0.5 Spearman")

cvData_0.9_S_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                         DecorPCAEFA=FALSE,thr=0.9,method="spearman",Outcome=outcome,drivingFeatures=outcome)
cvData_0.8_S_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                         DecorPCAEFA=FALSE,thr=0.8,method="spearman",Outcome=outcome,drivingFeatures=outcome)
cvData_0.7_S_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                         DecorPCAEFA=FALSE,thr=0.7,method="spearman",Outcome=outcome,drivingFeatures=outcome)
cvData_0.6_S_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                         DecorPCAEFA=FALSE,thr=0.6,method="spearman",Outcome=outcome,drivingFeatures=outcome)
cvData_0.5_S_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                         DecorPCAEFA=FALSE,thr=0.5,method="spearman",Outcome=outcome,drivingFeatures=outcome)
bpDe <- predictionStats_binary(cvData_0.9_S_D$DeMedTestRaw,"IDeA@0.9 Spearman Driven")
bpDe <- predictionStats_binary(cvData_0.8_S_D$DeMedTestRaw,"IDeA@0.8 Spearman Driven")
bpDe <- predictionStats_binary(cvData_0.7_S_D$DeMedTestRaw,"IDeA@0.7 Spearman Driven")
bpDe <- predictionStats_binary(cvData_0.6_S_D$DeMedTestRaw,"IDeA@0.6 Spearman Driven")
bpDe <- predictionStats_binary(cvData_0.5_S_D$DeMedTestRaw,"IDeA@0.5 Spearman Driven")


save.image(paste("ALLCV_analyiss",studyName,".RData",sep="_"))

```
