---
title: "Decorrelation-Based Feature Discovery: Seeds"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
  word_document: 
    reference_docx: WordStyle_FRESA.docx
    toc: yes
    fig_caption: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message = FALSE,comment = "#>")

```


# Effect of UPSTM-Based Decorrelation on Feature Discovery


### Loading the libraries

```{r}
library("FRESA.CAD")
library(readxl)
library(igraph)
library(umap)
library(tsne)
library(entropy)
library(psych)
library(whitening)
library("vioplot")
library("rpart")

op <- par(no.readonly = TRUE)
pander::panderOptions('digits', 3)
pander::panderOptions('table.split.table', 400)
pander::panderOptions('keep.trailing.zeros',TRUE)

```

## Material and Methods

Data Source 
https://archive.ics.uci.edu/ml/datasets/seeds


M. Charytanowicz, J. Niewczas, P. Kulczycki, P.A. Kowalski, S. Lukasik, S. Zak, 'A Complete Gradient Clustering Algorithm for Features Analysis of X-ray Images', in: Information Technologies in Biomedicine, Ewa Pietka, Jacek Kawa (eds.), Springer-Verlag, Berlin-Heidelberg, 2010, pp. 15-24.


## The Data

```{r}
seeds <- read.delim("~/GitHub/LatentBiomarkers/Data/seeds_dataset.txt", header=FALSE)
par(cex=0.5)

featnames <- c("area",
               "perimeter",
               "compactness",
               "length_of_kernel",
               "width_of_kernel",
               "asymmetry_coeff",
               "length_ker_groove",
               "class"
)
colnames(seeds) <- featnames
#seeds$class <- 1*(seeds$class == 1)
pander::pander(table(seeds$class))

```



## Comparing IDeA vs PCA vs EFA

### Getting the top correlated variables
```{r results = "asis" ,warning = FALSE, dpi=600, fig.height= 8, fig.width= 11}
plot(seeds[,colnames(seeds)!="class"],col=seeds$class,cex=0.65,cex.lab=0.5,cex.axis=0.75,cex.sub=0.5,cex.main=0.75)
rawCor <- cor(seeds[,colnames(seeds)!="class"])

  gplots::heatmap.2(abs(rawCor),
                    trace = "none",
  #                  scale = "row",
                    mar = c(5,5),
                    col=rev(heat.colors(5)),
                    main = "Original Correlation",
                    cexRow = 0.5,
                    cexCol = 0.5,
                     srtCol=45,
                     srtRow= -45,
                    key.title=NA,
                    key.xlab="Pearson Correlation",
                    xlab="Feature", ylab="Feature")



diag(rawCor) <- 0;
getMaxcor <- apply(rawCor,2,max)
topCorrelated <- getMaxcor[which.max(getMaxcor)]
whotoMax <- getMaxcor[getMaxcor == topCorrelated]
plot(seeds[,names(whotoMax)],main="Top Correlated variables")

```

### IDeA

```{r results = "asis" ,warning = FALSE, dpi=600, fig.height= 8, fig.width= 11}
IDeASeeds <- IDeA(seeds)
plot(IDeASeeds[,colnames(IDeASeeds)!="class"],col=seeds$class,cex=0.65,cex.lab=0.5,cex.axis=0.75,cex.sub=0.5,cex.main=0.75)
pander::pander(attr(IDeASeeds,"UPSTM"))
pander::pander(getLatentCoefficients(IDeASeeds))


IDeACor <- cor(IDeASeeds[,colnames(IDeASeeds)!="class"])


  gplots::heatmap.2(abs(IDeACor),
                    trace = "none",
  #                  scale = "row",
                    mar = c(5,5),
                    col=rev(heat.colors(5)),
                    main = "IDeA Correlation",
                    cexRow = 0.5,
                    cexCol = 0.5,
                     srtCol=45,
                     srtRow= -45,
                    key.title=NA,
                    key.xlab="Pearson Correlation",
                    xlab="Feature", ylab="Feature")


pander::pander(colnames(IDeACor))
diag(IDeACor) <- 0;
getMaxcor <- apply(IDeACor,2,max)
topCorrelated <- getMaxcor[which.max(getMaxcor)]
whotoMax <- getMaxcor[getMaxcor == topCorrelated]
plot(IDeASeeds[,names(whotoMax)],main="IDeA: Top Correlated variables")
plot(IDeASeeds[,c("area","La_perimeter")],main="IDeA: Top Raw Correlated variables")

```

### PCA

```{r results = "asis" ,warning = FALSE, dpi=600, fig.height= 8, fig.width= 11}
featuresnames <- colnames(seeds)[colnames(seeds) != "class"]
pc <- prcomp(seeds[,featuresnames],center = TRUE,tol=0.01)   #principal components
PCAseed <- as.data.frame(cbind(predict(pc,seeds[,featuresnames]),class=seeds$class))
plot(PCAseed[,colnames(PCAseed)!="class"],col=seeds$class,cex=0.65,cex.lab=0.5,cex.axis=0.75,cex.sub=0.5,cex.main=0.75)

pander::pander(pc$rotation)


PCACor <- cor(PCAseed[,colnames(PCAseed)!="class"])


  gplots::heatmap.2(abs(PCACor),
                    trace = "none",
  #                  scale = "row",
                    mar = c(5,5),
                    col=rev(heat.colors(5)),
                    main = "PCA Correlation",
                    cexRow = 0.5,
                    cexCol = 0.5,
                     srtCol=45,
                     srtRow= -45,
                    key.title=NA,
                    key.xlab="Pearson Correlation",
                    xlab="Feature", ylab="Feature")


```

### EFA


```{r results = "asis" ,warning = FALSE, dpi=600, fig.height= 8, fig.width= 11}
featuresnames <- colnames(seeds)[colnames(seeds) != "class"]
uls <- fa(seeds[,featuresnames],length(featuresnames)-1,rotate="varimax",warnings=FALSE)  # EFA analysis
EFAseed <- as.data.frame(cbind(predict(uls,seeds[,featuresnames]),class=seeds$class))
plot(EFAseed[,colnames(EFAseed)!="class"],col=seeds$class,cex=0.65,cex.lab=0.5,cex.axis=0.75,cex.sub=0.5,cex.main=0.75)
pander::pander(uls$weights)


EFACor <- cor(EFAseed[,colnames(EFAseed)!="class"])


  gplots::heatmap.2(abs(EFACor),
                    trace = "none",
  #                  scale = "row",
                    mar = c(5,5),
                    col=rev(heat.colors(5)),
                    main = "EFA Correlation",
                    cexRow = 0.5,
                    cexCol = 0.5,
                     srtCol=45,
                     srtRow= -45,
                    key.title=NA,
                    key.xlab="Pearson Correlation",
                    xlab="Feature", ylab="Feature")

```


## Effect on CAR modeling

```{r results = "asis", warning = FALSE, dpi=600, fig.height= 8, fig.width= 11}
par(op)
par(xpd = TRUE)
seeds$class <- factor(seeds$class)
rawmodel <- rpart(class~.,seeds,control=rpart.control(maxdepth=3))
plot(rawmodel,main="Raw",branch=0.5,uniform = TRUE,compress = TRUE,margin=0.1)
text(rawmodel, use.n = TRUE,cex=0.75)
pr <- predict(rawmodel,seeds,type = "class")
pander::pander(table(seeds$class,pr))
pander::pander(c(accuracy=sum(seeds$class==pr)/nrow(seeds)))

IDeASeeds$class <- factor(IDeASeeds$class)
IDeAmodel <- rpart(class~.,IDeASeeds,control=rpart.control(maxdepth=3))
plot(IDeAmodel,main="IDeA",branch=0.5,uniform = TRUE,compress = TRUE,margin=0.1)
text(IDeAmodel, use.n = TRUE,cex=0.75)
pr <- predict(IDeAmodel,IDeASeeds,type = "class")
pander::pander(table(IDeASeeds$class,pr))
pander::pander(c(accuracy=sum(IDeASeeds$class==pr)/nrow(seeds)))


PCAseed$class <- factor(PCAseed$class)
PCAmodel <- rpart(class~.,PCAseed,control=rpart.control(maxdepth=3))
plot(PCAmodel,main="PCA",branch=0.5,uniform = TRUE,compress = TRUE,margin=0.1)
text(PCAmodel, use.n = TRUE,cex=0.75)
pr <- predict(PCAmodel,PCAseed,type = "class")
pander::pander(table(PCAseed$class,pr))
pander::pander(c(accuracy=sum(PCAseed$class==pr)/nrow(seeds)))


EFAseed$class <- factor(EFAseed$class)
EFAmodel <- rpart(class~.,EFAseed,control=rpart.control(maxdepth=3))
plot(EFAmodel,main="EFA",branch=0.5,uniform = TRUE,compress = TRUE,margin=0.1)
text(EFAmodel, use.n = TRUE,cex=0.75)
pr <- predict(EFAmodel,EFAseed,type = "class")
pander::pander(table(EFAseed$class,pr))
pander::pander(c(accuracy=sum(EFAseed$class==pr)/nrow(seeds)))

par(op)
```


## Effect on logisitc modeling

```{r results = "asis", warning = FALSE, dpi=600, fig.height= 8, fig.width= 11}
seeds$class <- 1*(seeds$class == 1)
IDeASeeds$class <- seeds$class
PCAseed$class <- seeds$class
EFAseed$class <- seeds$class
rawmodel <- glm(class~.,seeds,family="binomial")
pander::pander(summary(rawmodel))
plot(rawmodel)

IDeAmodel <- glm(class~.,IDeASeeds,family="binomial")
pander::pander(summary(IDeAmodel))
plot(IDeAmodel)


PCAmodel <- glm(class~.,PCAseed,family="binomial")
pander::pander(summary(PCAmodel))
plot(PCAmodel)


EFAmodel <- glm(class~.,EFAseed,family="binomial")
pander::pander(summary(EFAmodel))
plot(EFAmodel)

par(op)
```



#### Standarize the names for the reporting

```{r results = "asis" }
studyName <- "Seeds"
dataframe <- seeds
outcome <- "class"
thro <- 0.80
TopVariables <- 5
cexheat = 0.45
```


## Generaring the report

```{r parent, child = here::here("~/GitHub/LatentBiomarkers/RMD/Child_univariate/child_Simple_LatentBiomarkers.Rmd")}
```


