---
title: "Decorrelation-Based Feature Discovery: Seeds"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
  word_document: 
    reference_docx: WordStyle_FRESA.docx
    toc: yes
    fig_caption: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message = FALSE,comment = "#>")

```

# Effect of UPSTM-Based Decorrelation on Feature Discovery



### Loading the libraries

```{r}
library("FRESA.CAD")
library(readxl)
library(igraph)
library(umap)
library(tsne)
library(entropy)

op <- par(no.readonly = TRUE)
pander::panderOptions('digits', 3)
pander::panderOptions('table.split.table', 400)
pander::panderOptions('keep.trailing.zeros',TRUE)
source("C:/Users/jtame/Documents/GitHub/LatentBiomarkers/RMD/DecorrelationCV.R")

```

## Material and Methods

Data Source 
https://archive.ics.uci.edu/ml/datasets/seeds


M. Charytanowicz, J. Niewczas, P. Kulczycki, P.A. Kowalski, S. Lukasik, S. Zak, 'A Complete Gradient Clustering Algorithm for Features Analysis of X-ray Images', in: Information Technologies in Biomedicine, Ewa Pietka, Jacek Kawa (eds.), Springer-Verlag, Berlin-Heidelberg, 2010, pp. 15-24.


## The Data

```{r}
seeds <- read.delim("~/GitHub/LatentBiomarkers/Data/seeds_dataset.txt", header=FALSE)

featnames <- c("area",
               "perimeter",
               "compactness",
               "length_of_kernel",
               "width_of_kernel",
               "asymmetry_coeff",
               "length_ker_groove",
               "class"
)
colnames(seeds) <- featnames
seeds$class <- 1*(seeds$class == 1)
pander::pander(table(seeds$class))

```


#### Standarize the names for the reporting

```{r results = "asis"}
studyName <- "Seeds"
dataframe <- seeds
outcome <- "class"
nLoops <- 100

needScaling <- TRUE
IDSample <- FALSE
thro <- 0.8

```
## CV Analysis and report

```{r parent, child = here::here("~/GitHub/LatentBiomarkers/RMD/Child_univariate/CV_analysis.Rmd")}
```



## Testing different parameters
```{r}
nLoops <- 30

cvData_0.9 <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.9)
cvData_0.8 <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.8)
cvData_0.7 <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.7)
cvData_0.6 <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.6)
cvData_0.5 <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.5)
bpDe <- predictionStats_binary(cvData_0.9$DeMedTestRaw,"ILAA@0.9")
bpDe <- predictionStats_binary(cvData_0.8$DeMedTestRaw,"ILAA@0.8")
bpDe <- predictionStats_binary(cvData_0.7$DeMedTestRaw,"ILAA@0.7")
bpDe <- predictionStats_binary(cvData_0.6$DeMedTestRaw,"ILAA@0.6")
bpDe <- predictionStats_binary(cvData_0.5$DeMedTestRaw,"ILAA@0.5")


cvData_0.9_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                       DecorPCAEFA=FALSE,thr=0.9,Outcome=outcome,drivingFeatures=outcome)
cvData_0.8_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                       DecorPCAEFA=FALSE,thr=0.8,Outcome=outcome,drivingFeatures=outcome)
cvData_0.7_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                       DecorPCAEFA=FALSE,thr=0.7,Outcome=outcome,drivingFeatures=outcome)
cvData_0.6_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                       DecorPCAEFA=FALSE,thr=0.6,Outcome=outcome,drivingFeatures=outcome)
cvData_0.5_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                       DecorPCAEFA=FALSE,thr=0.5,Outcome=outcome,drivingFeatures=outcome)
bpDe <- predictionStats_binary(cvData_0.9_D$DeMedTestRaw,"ILAA@0.9 Driven")
bpDe <- predictionStats_binary(cvData_0.8_D$DeMedTestRaw,"ILAA@0.8 Driven")
bpDe <- predictionStats_binary(cvData_0.7_D$DeMedTestRaw,"ILAA@0.7 Driven")
bpDe <- predictionStats_binary(cvData_0.6_D$DeMedTestRaw,"ILAA@0.6 Driven")
bpDe <- predictionStats_binary(cvData_0.5_D$DeMedTestRaw,"ILAA@0.5 Driven")

cvData_0.9_S <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.9,method="spearman",type="RLM")
cvData_0.8_S <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.8,method="spearman",type="RLM")
cvData_0.7_S <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.7,method="spearman",type="RLM")
cvData_0.6_S <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.6,method="spearman",type="RLM")
cvData_0.5_S <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,DecorPCAEFA=FALSE,thr=0.5,method="spearman",type="RLM")
bpDe <- predictionStats_binary(cvData_0.9_S$DeMedTestRaw,"ILAA@0.9 Spearman")
bpDe <- predictionStats_binary(cvData_0.8_S$DeMedTestRaw,"ILAA@0.8 Spearman")
bpDe <- predictionStats_binary(cvData_0.7_S$DeMedTestRaw,"ILAA@0.7 Spearman")
bpDe <- predictionStats_binary(cvData_0.6_S$DeMedTestRaw,"ILAA@0.6 Spearman")
bpDe <- predictionStats_binary(cvData_0.5_S$DeMedTestRaw,"ILAA@0.5 Spearman")

cvData_0.9_S_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                         DecorPCAEFA=FALSE,thr=0.9,method="spearman",Outcome=outcome,drivingFeatures=outcome,type="RLM")
cvData_0.8_S_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                         DecorPCAEFA=FALSE,thr=0.8,method="spearman",Outcome=outcome,drivingFeatures=outcome,type="RLM")
cvData_0.7_S_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                         DecorPCAEFA=FALSE,thr=0.7,method="spearman",Outcome=outcome,drivingFeatures=outcome,type="RLM")
cvData_0.6_S_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                         DecorPCAEFA=FALSE,thr=0.6,method="spearman",Outcome=outcome,drivingFeatures=outcome,type="RLM")
cvData_0.5_S_D <- CV_TDR(dataframe,outcome,loops=nLoops,IDSample=IDSample,scale=FALSE,
                         DecorPCAEFA=FALSE,thr=0.5,method="spearman",Outcome=outcome,drivingFeatures=outcome,type="RLM")
bpDe <- predictionStats_binary(cvData_0.9_S_D$DeMedTestRaw,"ILAA@0.9 Spearman Driven")
bpDe <- predictionStats_binary(cvData_0.8_S_D$DeMedTestRaw,"ILAA@0.8 Spearman Driven")
bpDe <- predictionStats_binary(cvData_0.7_S_D$DeMedTestRaw,"ILAA@0.7 Spearman Driven")
bpDe <- predictionStats_binary(cvData_0.6_S_D$DeMedTestRaw,"ILAA@0.6 Spearman Driven")
bpDe <- predictionStats_binary(cvData_0.5_S_D$DeMedTestRaw,"ILAA@0.5 Spearman Driven")


save.image(paste("ALLCV_analyiss",studyName,".RData",sep="_"))

```


