---
title: "Decorrelation-Based SOTA: Seeds"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
  word_document: 
    reference_docx: WordStyle_FRESA.docx
    toc: yes
    fig_caption: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message = FALSE,comment = "#>")

```

# Effect of UPSTM-Based Decorrelation on ML Performance



### Loading the libraries

```{r}
library("FRESA.CAD")
library(readxl)
library(igraph)
library(umap)
library(tsne)
library(entropy)

op <- par(no.readonly = TRUE)
pander::panderOptions('digits', 3)
pander::panderOptions('table.split.table', 400)
pander::panderOptions('keep.trailing.zeros',TRUE)

```

## Material and Methods

Data Source 
https://archive.ics.uci.edu/ml/datasets/seeds


M. Charytanowicz, J. Niewczas, P. Kulczycki, P.A. Kowalski, S. Lukasik, S. Zak, 'A Complete Gradient Clustering Algorithm for Features Analysis of X-ray Images', in: Information Technologies in Biomedicine, Ewa Pietka, Jacek Kawa (eds.), Springer-Verlag, Berlin-Heidelberg, 2010, pp. 15-24.


## The Data

```{r}
seeds <- read.delim("~/GitHub/LatentBiomarkers/Data/seeds_dataset.txt", header=FALSE)

seeds$V8
featnames <- c("area",
               "perimeter",
               "compactness",
               "length_of_kernel",
               "width_of_kernel",
               "asymmetry_coeff",
               "length_ker_groove",
               "class"
)
colnames(seeds) <- featnames
seeds$class <- 1*(seeds$class == 1)
pander::pander(table(seeds$class))

```


#### Standarize the names for the reporting

```{r results = "asis"}
studyName <- "Seeds"
dataframe <- seeds
outcome <- "class"
nLoops <- 50
tfrac <- 0.80

needScaling <- TRUE
IDSample <- FALSE
thro <- 0.8


```
## CV Analysis and report

```{r parent, child = here::here("~/GitHub/LatentBiomarkers/RMD/Child_univariate/SOTACV_analysis.Rmd")}
```



