---
title: "Decorrelation-Based Feature Discovery: Echo"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
  word_document: 
    reference_docx: WordStyle_FRESA.docx
    toc: yes
    fig_caption: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE, message = FALSE,comment = "#>")

```


# Effect of UPSTM-Based Decorrelation on Feature Discovery


### Loading the libraries

```{r}
library("FRESA.CAD")
library(readxl)
library(igraph)
library(umap)
library(tsne)
library(entropy)

op <- par(no.readonly = TRUE)
pander::panderOptions('digits', 3)
pander::panderOptions('table.split.table', 400)
pander::panderOptions('keep.trailing.zeros',TRUE)

```

## Material and Methods

Data Source 
https://archive.ics.uci.edu/ml/datasets/Echocardiogram

Salzberg, S. (1988). Exemplar-based learning: Theory and implementation (Technical Report TR-10-88). Harvard University, Center for Research in Computing Technology, Aiken Computation Laboratory (33 Oxford Street; Cambridge, MA 02138).
[Web Link]

Kan, G., Visser, C., Kooler, J., & Dunning, A. (1986). Short and long term predictive value of wall motion score in acute myocardial infarction. British Heart Journal, 56, 422-427.


## The Data

```{r}
Echodata <- read.csv("~/GitHub/LatentBiomarkers/Data/Echo/echocardiogram.data", header=FALSE, stringsAsFactors=TRUE)
featNames <- read.csv("~/GitHub/LatentBiomarkers/Data/Echo/names.csv")
colnames(Echodata) <- featNames$Name
Echodata$name <- NULL
Echodata$group <- NULL
Echodata$mult <- NULL
Echodata$alive_at_1 <- NULL
Echodata$survival <- NULL
Echodata[Echodata=="?"] <- NA
Echodata[,1:ncol(Echodata)] <- sapply(Echodata,as.numeric)
Echodata <- Echodata[complete.cases(Echodata),]
Echodata$still_alive <- 1*(Echodata$still_alive==3)
pander::pander(table(Echodata$still_alive))
```


#### Standarize the names for the reporting

```{r results = "asis"}
studyName <- "Echo"
dataframe <- Echodata
outcome <- "still_alive"

TopVariables <- 3

thro <- 0.80
cexheat = 0.15

```


## Generaring the report

```{r parent, child = here::here("~/GitHub/LatentBiomarkers/RMD/Child_univariate/child_Simple_LatentBiomarkers.Rmd")}
```